---
title: "Bamboo mesocosms experiment - squeezemeta analysis"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Setup

```{r}
library("SQMtools")
library(data.table)
library("phyloseq")
library("ggplot2")
```

# Individual sqm objects

Just to check stuff out

```{r, eval=F}
setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/metagenomes/SQM_zips_3samps")

lb5 <- loadSQM("LB5.zip")
wb9 <- loadSQM("WB9.zip")
ll11 <- loadSQM("LL11.zip")

# plotTaxonomy(lb5)
# plotTaxonomy(wb9)
# plotTaxonomy(ll11)

alltab.3samps <- combineSQMlite(lb5,wb9,ll11)

alltab.tax <- alltab.3samps[["taxa"]]

##Just bacteria:
lb5.bac = subsetTax(lb5, "superkingdom", "Bacteria")
wb9.bac = subsetTax(wb9, "superkingdom", "Bacteria")
ll11.bac = subsetTax(ll11, "superkingdom", "Bacteria")

bactab.3samps <- combineSQMlite(lb5.bac,wb9.bac,ll11.bac)
```

# Making a big taxa counts table for all samples & all taxa

On the cluster, text within alltax.R:

```{r, eval=F}
library("SQMtools")

setwd("/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/mostsamps/rready/all_proc")
#setwd("/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/3samps/rready/allproc")

zip_dir <- "/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/mostsamps/rready"
#zip_dir <- "/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/3samps/rready"

# Get a list of all .zip files
zip_files <- list.files(zip_dir, pattern = "\\.zip$", full.names = TRUE)
head(zip_files)

# Loop through each .zip file
for (zip_file in zip_files){
  # Extract the base name (without extension) to use as the object name
  obj_name <- tools::file_path_sans_ext(basename(zip_file))
  # Assign the result of loadSQM() to a variable with the extracted name
  assign(obj_name, loadSQM(zip_file), envir = .GlobalEnv)
}

##selecting my R objects that start with these 2 capital letters
object_names <- ls(pattern = "^[LW]")
object_names

# Retrieve the actual objects using mget()
object_list <- mget(object_names)

# Combine them into one using combineSQMlite()
allsamps <- do.call(combineSQMlite, object_list)

saveRDS(allsamps, file="allsamps_alltax.rds")
```

Batch job

```{bash, eval=F}
nano alltax.slurm
```

Text within

```
#!/bin/bash
#SBATCH --job-name=alltax
#SBATCH --partition=shared
##3 day max run time for public partitions, except 4 hour max runtime for the s>
#SBATCH --time=0-24:00:00
#SBATCH --nodes=1
#SBATCH --cpus-per-task=10
#SBATCH --mem=100G
#SBATCH --error=%A.err
#SBATCH --output=%A.out ##%A = filled with job id
#SBATCH --mail-type=BEGIN,END,FAIL,REQUEUE,TIME_LIMIT_80
#SBATCH --mail-user=nicolagk@hawaii.edu

module load lang/R/4.2.1-foss-2022a
Rscript alltax.R
```

```{bash, eval=F}
sbatch alltax.slurm
```

# Basic info

```{r}
setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/06.metagnomes")

alltab <- readRDS("allsamps_alltax.rds")

fxn.counts <- data.frame(alltab[["functions"]][["KEGG"]][["abund"]])
dim(fxn.counts) #13362 functions
length(unique(row.names(fxn.counts)))

fxn.names <- data.frame(gene_names=alltab[["KEGG_names"]])
#12451.... where are the rest? I assume de-duplicated

fxn.mismatch <- fxn.counts[!row.names(fxn.counts) %in% row.names(fxn.names),]

alltab.fxn <- alltab[["misc"]]
fxn.names <- data.frame(gene_names=alltab.fxn[["KEGG_names"]])

##
all.counts.sp <- data.frame(alltab.fxn[["species"]][["abund"]])
#4869 taxa, 48 samples 
```

# Retrieving all kegg info separately

I think when the sqm objects are combined, the duplicate gene names are removed, but they do have unique kegg ids, so I'm losing information because the counts table is not de-duplicated. Going to grab all the .sqm objects separately and make one big table with all the kegg info.

Text within alltax_kegg.R:

```{r, eval=F}
library("SQMtools")

setwd("/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/mostsamps/rready/all_proc")
#setwd("/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/3samps/rready/allproc")
#setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/metagenomes/SQM_zips_3samps")

#zip_dir <- "~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/metagenomes/SQM_zips_3samps"
zip_dir <- "/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/mostsamps/rready"
# zip_dir <- "/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/3samps/rready"

# Get a list of all .zip files
zip_files <- list.files(zip_dir, pattern = "\\.zip$", full.names = TRUE)
head(zip_files)

# Create an empty list to store data frames
fxn_list <- list()

# Loop through each .zip file
for (zip_file in zip_files) {
  # Extract the base name (without extension) to use as the object name
  obj_name <- tools::file_path_sans_ext(basename(zip_file))
  
  # Load the SQM object
  sqm_obj <- loadSQM(zip_file)
  
  # Pull out the functional annotations
  alltab_fxn <- sqm_obj[["misc"]]
  
  # Create a data frame extracting available KEGG info
  fxn_names <- data.frame(
    #sample = obj_name,
    kegg_id = names(alltab_fxn[["KEGG_names"]]),
    gene_name = alltab_fxn[["KEGG_names"]],
    gene_path = alltab_fxn[["KEGG_paths"]]
    #stringsAsFactors = FALSE
  )
  
  # Store each data frame in the list
  #fxn_list[[obj_name]] <- fxn_names
  # Save each data frame into the global environment
  assign(paste0(obj_name, "_fxndf"), fxn_names, envir = .GlobalEnv)
}

# Get all objects whose names end in "_fxndf"
fxn_dfs <- mget(ls(pattern = "_fxndf"))

stacked_df <- do.call(rbind, fxn_dfs)

stacked_df2 <- stacked_df[!duplicated(stacked_df$kegg_id), ]

write.csv(stacked_df2,file="allsamps_alltax_FULLkegginfo.csv")
```

Batch job, text within "alltax_kegg.slurm":

```
#!/bin/bash
#SBATCH --job-name=alltax_kegg
#SBATCH --partition=shared
##3 day max run time for public partitions, except 4 hour max runtime for the s>
#SBATCH --time=0-24:00:00
#SBATCH --nodes=1
#SBATCH --cpus-per-task=10
#SBATCH --mem=100G
#SBATCH --error=%A.err
#SBATCH --output=%A.out ##%A = filled with job id
#SBATCH --mail-type=BEGIN,END,FAIL,REQUEUE,TIME_LIMIT_80
#SBATCH --mail-user=nicolagk@hawaii.edu

module load lang/R/4.2.1-foss-2022a
Rscript alltax_kegg.R
```

Submit

```{bash, eval=F}
sbatch alltax_kegg.slurm
```

That worked! Now I have the kegg ids, gene names, and pathways for 13k functions

# Checking out taxa tables in R

## Making a counts table

Down to species I guess, I don't see the ORF ID as the unit

```{r}
setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/06.metagnomes")

alltab <- readRDS("allsamps_alltax.rds")

alltab.tax <- alltab[["taxa"]]

##counts table
all.counts.sp <- data.frame(alltab.tax[["species"]][["abund"]])
#4869 taxa, 48 samples 
```

## Taxa table

```{r}
##taxa table 
all.tax.names <- data.table(full_names=alltab[["misc"]][["tax_names_long"]][["species"]])

# Split into multiple columns
all.tax.names.exp <- all.tax.names[,tstrsplit(full_names, ";", fixed = TRUE)]
head(all.tax.names.exp)
##removing the first letter and the underscore
all.tax.names.exp[, (names(all.tax.names.exp)) := lapply(.SD, function(x) gsub("^[a-zA-Z]_", "", x))]

all.tax.names.exp <- data.frame(all.tax.names.exp)
colnames(all.tax.names.exp) <- names(alltab.tax)
row.names(all.tax.names.exp) <- all.tax.names.exp$species

any(!row.names(all.tax.names.exp) %in% row.names(all.counts.sp))
```

## Sample data

```{r}
metadat <- read.csv("../bamboomesos_metadata_plusmgs.csv")
row.names(metadat) <- metadat$Short_label

##sample type to glom by
#metadat$sam_glom <- paste0()
```

## Phyloseq object

```{r}
ps <- phyloseq(otu_table(all.counts.sp,taxa_are_rows=T),
         sample_data(metadat),
         tax_table(as.matrix(all.tax.names.exp)))
ps
# otu_table()   OTU Table:         [ 4869 taxa and 48 samples ]
# sample_data() Sample Data:       [ 48 samples by 25 sample variables ]
# tax_table()   Taxonomy Table:    [ 4869 taxa by 7 taxonomic ranks ]


#ps.sampsum <- merge_samples(ps, "")

ps.rel <- transform_sample_counts(ps, function(x) x / sum(x))
ps.rel.king <- tax_glom(ps.rel,"superkingdom")

plot_bar(ps.rel.king,fill="superkingdom")

plot_bar(ps.rel.king,fill="superkingdom")+
  facet_wrap(~superkingdom,scales="free")

ps.rel.euk <- subset_taxa(ps.rel,superkingdom=="Eukaryota")

plot_bar(ps.rel.euk,fill="phylum")
```

# Subsetting taxa-specific functions

## Bacteria functions

```{r, eval=F}
library("SQMtools")

setwd("/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/mostsamps/rready/rproc")

##real deal:
zip_dir <- "/home/nicolagk/cmaiki_koastore/nicolagk/bamboomesos_metagnomes/mostsamps/rready"

# Get a list of all .zip files
zip_files <- list.files(zip_dir, pattern = "\\.zip$", full.names = TRUE)
head(zip_files)

# Loop through each .zip file
for (zip_file in zip_files){
  # Extract the base name (without extension) to use as the object name
  obj_name <- tools::file_path_sans_ext(basename(zip_file))
  # Assign the result of loadSQM() to a variable with the extracted name
  assign(obj_name, loadSQM(zip_file), envir = .GlobalEnv)
}

##selecting my R objects that start with these 2 capital letters
object_names <- ls(pattern = "^[LW]")

for (obj in object_names) {
  # Construct the new variable name
  new_name <- paste0(obj, ".bac")
  # Print the name for debugging
  print(paste("Creating:", new_name))
  # Use get() to retrieve the actual object
  obj_data <- get(obj)
  # Apply the function to the object
  result <- subsetTax(obj_data, "superkingdom", "Bacteria")
  # Assign the result to a new variable with the dynamically created name
  assign(new_name, result, envir = .GlobalEnv)
}

# Get all objects that end with ".bac" (matching our dynamically created ones)
bac_objects <- ls(pattern = "\\.bac$")

# Retrieve the actual objects using mget()
bac_list <- mget(bac_objects)

# Combine them into one using combineSQMlite()
bactab.all <- do.call(combineSQMlite, bac_list)

saveRDS(bactab.all,file="./bactab.all.rds")
```
