---
title: "Metagenomes abundance analyses"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Setup

## Libs

```{r}
#install.packages("remotes")
#install.packages('TMB',source=TRUE) #didn't work on home laptop
#remotes::install_github("glmmTMB/glmmTMB/glmmTMB")
library("glmmTMB")
#install.packages("BiocManager")
#BiocManager::install("phyloseq")
library("phyloseq")
library("dplyr")
# # #install.packages("reshape")
# library("reshape")
# # library("ggplot2")
# # #install.packages("bbmle")
library("bbmle")
# # #install.packages("car")
# # library("car")
# library("DHARMa")
# library("plyr")
library("ggplot2")
library("data.table")
```

## Re-read data

```{r}
setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/06.metagnomes/")

bm.metadat <- read.csv("../bamboomesos_metadata_plusmgs.csv")
row.names(bm.metadat) <- bm.metadat$Short_label

bm.bactab1 <- readRDS("bactab.all.rds")
bm.bactab <- bm.bactab1[["functions"]][["KEGG"]][["abund"]]

bm.kegginfo <- read.csv("allsamps_alltax_FULLkegginfo.csv",row.names=1)
row.names(bm.kegginfo) <- bm.kegginfo$kegg_id

#split the pathways column into multiple columns

#strsplit(bm.kegginfo$Pathway, "; ", fixed = TRUE)
# bm.kegginfo$kegg.path <- paste0(bm.kegginfo$KO,"; ", bm.kegginfo$Pathway)
# bm.kegginfo.dt <- data.table(bm.kegginfo)
# bm.kegginfo.expanded <- bm.kegginfo.dt[,tstrsplit(kegg.path, ";", fixed = TRUE)]
#colnames(bm.kegginfo)[1] <- c("KO")
#bm.kegginfo2 <- merge(bm.kegginfo.dt,bm.kegginfo.expanded,by="KO")
```

# Set up for mixed mods

## Make phyloseq object

Not really necessary, but I like how phyloseq kicks out all of the irrelevant data at once

```{r}
ps.fxn <- phyloseq(otu_table(bm.bactab,taxa_are_rows=T),
         tax_table(as.matrix(bm.kegginfo)),
         sample_data(bm.metadat))

ps.fxn #9072 taxa, 48 samples
```

## Split by sample types

```{r}
ps.fxn.lar <- subset_samples(ps.fxn,Sample_type=="Larvae")
ps.fxn.lar ##25 samples

ps.fxn.wat <- subset_samples(ps.fxn,Sample_type=="Water")
ps.fxn.wat ##23 samples
```

## Filter

```{r}
##Matt did 12%, so 5.76 samples out of 48
##I'll be doing:
#6/25 #0.24 for larvae
#6/23 #0.26 for water

##larvae
ps.fxn.lar.trim <- filter_taxa(ps.fxn.lar, function(x) sum(x > 1) > (0.24*length(x)), TRUE)
ps.fxn.lar.trim ##5748 taxa
tail(sort(sample_sums(ps.fxn.lar.trim),decreasing=T))

##water
ps.fxn.wat.trim <- filter_taxa(ps.fxn.wat, function(x) sum(x > 1) > (0.26*length(x)), TRUE)
ps.fxn.wat.trim ##7186 taxa
tail(sort(sample_sums(ps.fxn.wat.trim),decreasing=T))
```

## Check zeroes levels

```{r}
##larvae - after trimming
otu.fxn.lar.trim <- data.frame(ps.fxn.lar.trim@otu_table)
sum(otu.fxn.lar.trim == 0) / (ncol(otu.fxn.lar.trim) * nrow(otu.fxn.lar.trim)) * 100
#18.63535%

##without trimming
otu.fxn.lar <- data.frame(ps.fxn.lar@otu_table)
sum(otu.fxn.lar == 0) / (ncol(otu.fxn.lar) * nrow(otu.fxn.lar)) * 100
#46.17063%

##water - after trimming
otu.fxn.wat.trim <- data.frame(ps.fxn.wat.trim@otu_table)
sum(otu.fxn.wat.trim == 0) / (ncol(otu.fxn.wat.trim) * nrow(otu.fxn.wat.trim)) * 100
#14.39454%

##without trimming
otu.fxn.wat <- data.frame(ps.fxn.wat@otu_table)
sum(otu.fxn.wat == 0) / (ncol(otu.fxn.wat) * nrow(otu.fxn.wat)) * 100
#30.55556%

##save if ready
#saveRDS(ps.fxn.lar.trim,file="ps.fxn.lar.trim.rds")
#saveRDS(ps.fxn.wat.trim,file="ps.fxn.wat.trim.rds")
```

## More formatting - long form

```{r}
otu.wat <- data.frame(t(ps.fxn.wat.trim@otu_table))
otu.lar <- data.frame(t(ps.fxn.lar.trim@otu_table))

#total read count colums
otu.wat$sum <- rowSums(otu.wat)
otu.lar$sum <- rowSums(otu.lar)

otu.wat$Short_label <- row.names(otu.wat)
otu.lar$Short_label <- row.names(otu.lar)

#bring in the metadata again
samdf.wat <- data.frame(ps.fxn.wat.trim@sam_data)
samdf.lar <- data.frame(ps.fxn.lar.trim@sam_data)

# #join the data sets by sample name
df.water <- plyr::join(otu.wat, samdf.wat, by="Short_label", type="left")
df.larv <- plyr::join(otu.lar, samdf.lar, by="Short_label", type="left")

##getting rid of some irrelevant metadat
#colnames(df.water)
df.water1 <- df.water %>%
  dplyr::select(-Short_label,-Longer_name,-Sample_type,-Num_larvae,-Date_collected,-Exp_day,-Mesocosm_treatment,-Mesocosm_type,-Treatment_notes,-Notes,-Num_larvae_d00,-Num_larvae_d08,-Num_larvae_d34,-Num_larvae_d40,-Total_pupae,-Proportion_pupated,-Biomass_day30,-metagenomes,-SSU_raw_reads,-X16S_raw_reads,-MG_raw_reads)
#colnames(df.water1)

df.water.long <- reshape2::melt(df.water1, id.vars=c("sum", "Mesocosm_id", "Microbe_treatment","Food_type", "Biomass_day40"))

df.water.long$rowid <- 1:nrow(df.water.long)
##165278 rows

df.larv1 <- df.larv %>%
  dplyr::select(-Short_label,-Longer_name,-Sample_type,-Num_larvae,-Date_collected,-Exp_day,-Mesocosm_treatment,-Mesocosm_type,-Treatment_notes,-Notes,-Num_larvae_d00,-Num_larvae_d08,-Num_larvae_d34,-Num_larvae_d40,-Total_pupae,-Proportion_pupated,-Biomass_day30,-metagenomes,-SSU_raw_reads,-X16S_raw_reads,-MG_raw_reads)
#colnames(df.larv1)

df.larv.long <- reshape2::melt(df.larv1, id.vars=c("sum", "Mesocosm_id", "Microbe_treatment","Food_type", "Biomass_day40"))

df.larv.long$rowid <- 1:nrow(df.larv.long)
##143700 rows

##save
#saveRDS(df.water.long,file="df.water.long.rds")
#saveRDS(df.larv.long,file="df.larv.long.rds")
```

# Running the statz{.tabset}

Can start from here

```{r}
setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/06.metagnomes")

df.water.long <- readRDS("df.water.long.rds")
df.larv.long <- readRDS("df.larv.long.rds")
```

## Water

Each model took around 10 minutes, ran them as background jobs "justmods_water.R" and "justmods_lar.R"

```
library("glmmTMB")
library("bbmle")

setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/06.metagnomes")

df.water.long <- readRDS("df.water.long.rds")
#df.larv.long <- readRDS("df.larv.long.rds")

str(df.water.long)
df.water.long$rowid <- as.factor(df.water.long$rowid)
df.water.long$Microbe_treatment <- as.factor(df.water.long$Microbe_treatment)
df.water.long$Food_type <- as.factor(df.water.long$Food_type)
#df.water.long$food.microbes <- as.factor(df.water.long$food.microbes)
df.water.long$Mesocosm_id <- as.factor(df.water.long$Mesocosm_id)
str(df.water.long)

water.mod.nb1 <- glmmTMB(value~offset(log(sum))+Microbe_treatment+Food_type+(1|variable)+(1|variable:Microbe_treatment)+(1|variable:Food_type),family=nbinom1,data=df.water.long)
summary(water.mod.nb1)

water.mod.nb2 <- glmmTMB(value~offset(log(sum))+Microbe_treatment+Food_type+(1|variable)+(1|variable:Microbe_treatment)+(1|variable:Food_type),family=nbinom2,data=df.water.long)
summary(water.mod.nb2)

AICtab(water.mod.nb1,water.mod.nb2)

saveRDS(water.mod.nb1,file="water.mod.nb1.full.rds")
saveRDS(water.mod.nb2,file="water.mod.nb2.full.rds")

#water.mod.nb1 <- readRDS("water.mod.nb1.full.rds")

water.mod.nb1.nomicr <- update(water.mod.nb1,.~.-(1|variable:Microbe_treatment))

#anova(water.mod.nb1,water.mod.nb1.nomic)

water.mod.nb1.nofood <- update(water.mod.nb1,.~.-(1|variable:Food_type))

#anova(water.mod.nb1,water.mod.nb1.nomic)

saveRDS(water.mod.nb1.nomicr,file="water.mod.nb1.nomicrobes.rds")
saveRDS(water.mod.nb1.nofood,file="water.mod.nb1.nofood.rds")
```

### Assessing models

```{r}
setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/06.metagnomes")

mod.water.nb2 <- readRDS("water.mod.nb2.full.rds")
mod.water.nb1 <- readRDS("water.mod.nb1.full.rds")
summary(mod.water.nb2)
summary(mod.water.nb1)

AICtab(mod.water.nb1,mod.water.nb2)
##nbinom1 better as usual

##reduced models
mod.water.nb1.nomicr <- readRDS("water.mod.nb1.nomicrobes.rds")
summary(mod.water.nb1.nomicr)

anova(mod.water.nb1.nomicr,mod.water.nb1) ##2.2e-16

mod.water.nb1.nofood <- readRDS("water.mod.nb1.nofood.rds")
summary(mod.water.nb1.nofood)

anova(mod.water.nb1.nofood,mod.water.nb1) ##2.2e-16

###save conditional modes
water.mod.ranef <- as.data.frame(ranef(mod.water.nb1))
#43116 rows

#write.csv(water.mod.ranef, "water.mod.ranef.csv")
```

## Larvae

```
library("glmmTMB")
library("bbmle")

setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/06.metagnomes")

#df.larv.long <- readRDS("df.larv.long.rds")
df.larv.long <- readRDS("df.larv.long.rds")

str(df.larv.long)
df.larv.long$rowid <- as.factor(df.larv.long$rowid)
df.larv.long$Microbe_treatment <- as.factor(df.larv.long$Microbe_treatment)
df.larv.long$Food_type <- as.factor(df.larv.long$Food_type)
#df.larv.long$food.microbes <- as.factor(df.larv.long$food.microbes)
df.larv.long$Mesocosm_id <- as.factor(df.larv.long$Mesocosm_id)
str(df.larv.long)

larv.mod.nb1 <- glmmTMB(value~offset(log(sum))+Microbe_treatment+Food_type+(1|variable)+(1|variable:Microbe_treatment)+(1|variable:Food_type),family=nbinom1,data=df.larv.long)
summary(larv.mod.nb1)

larv.mod.nb2 <- glmmTMB(value~offset(log(sum))+Microbe_treatment+Food_type+(1|variable)+(1|variable:Microbe_treatment)+(1|variable:Food_type),family=nbinom2,data=df.larv.long)
summary(larv.mod.nb2)

AICtab(larv.mod.nb1,larv.mod.nb2)

saveRDS(larv.mod.nb1,file="larv.mod.nb1.full.rds")
saveRDS(larv.mod.nb2,file="larv.mod.nb2.full.rds")

larv.mod.nb1 <- readRDS("larv.mod.nb1.full.rds")

larv.mod.nb1.nomicr <- update(larv.mod.nb1,.~.-(1|variable:Microbe_treatment))

anova(larv.mod.nb1,larv.mod.nb1.nomic)

larv.mod.nb1.nofood <- update(larv.mod.nb1,.~.-(1|variable:Food_type))

anova(larv.mod.nb1,larv.mod.nb1.nomic)

saveRDS(larv.mod.nb1.nomicr,file="larv.mod.nb1.nomicrobes.rds")
saveRDS(larv.mod.nb1.nofood,file="larv.mod.nb1.nofood.rds")
```

### Assessing models

```{r}
mod.larv.nb2 <- readRDS("larv.mod.nb2.full.rds")
mod.larv.nb1 <- readRDS("larv.mod.nb1.full.rds")

summary(mod.larv.nb2)
summary(mod.larv.nb1)

AICtab(mod.larv.nb1,mod.larv.nb2)
##nbinom1 better as usual

mod.larv.nb1.nomicr <- readRDS("larv.mod.nb1.nomicrobes.rds")
mod.larv.nb1.nofood <- readRDS("larv.mod.nb1.nofood.rds")

##no microbes
anova(mod.larv.nb1,mod.larv.nb1.nomicr) ##2.2e-16
anova(mod.larv.nb1,mod.larv.nb1.nofood) ##2.2e-16

##save conditional modes
larv.mod.ranef <- as.data.frame(ranef(mod.larv.nb1))
#write.csv(larv.mod.ranef, "larv.mod.ranef.csv")
#34488 rows
```

# Plotting

## Water

```{r}
library("stringr")
library("tidyr")

ps.fxn.wat.trim <- readRDS("ps.fxn.wat.trim.rds")

table(water.mod.ranef$grpvar)

water.ranef.food <- subset(water.mod.ranef, grpvar=="variable:Food_type")
water.ranef.micr <- subset(water.mod.ranef, grpvar=="variable:Microbe_treatment")
table(water.ranef.food$grpvar)
table(water.ranef.micr$grpvar)

water.ranef.food$id <- str_split_fixed(water.ranef.food$grp,":",n=2)[,1]
water.ranef.food$food <- str_split_fixed(water.ranef.food$grp,":",n=2)[,2]

water.ranef.micr$id <- str_split_fixed(water.ranef.micr$grp,":",n=2)[,1]
water.ranef.micr$microbes <- str_split_fixed(water.ranef.micr$grp,":",n=2)[,2]

##just the columns for stuff below:
water.ranef.food.less <- water.ranef.food %>%
  select(-component,-term,-grp,-condsd)

water.ranef.micr.less <- water.ranef.micr %>%
  select(-component,-term,-grp,-condsd)

# Reshape the data frame to have the food types as two columns
water.ranef.food.wide <- water.ranef.food.less %>%
  pivot_wider(
    names_from = food,   # Column to use for new column names
    values_from = condval       # Column to fill the new columns
  )

water.ranef.micr.wide <- water.ranef.micr.less %>%
  pivot_wider(
    names_from = microbes,   # Column to use for new column names
    values_from = condval       # Column to fill the new columns
  )

##add kegg info
fxn.water.trim <- as.data.frame(ps.fxn.wat.trim@tax_table)
fxn.water.trim$id <- fxn.water.trim$KO

water.ranef.micr.wide.tax <- merge(water.ranef.micr.wide,fxn.water.trim,by="id")
water.ranef.food.wide.tax <- merge(water.ranef.food.wide,fxn.water.trim,by="id")
##more informative kegg ids
water.ranef.micr.wide.tax$kegg.name <- paste0(water.ranef.micr.wide.tax$id,"_",water.ranef.micr.wide.tax$names)
water.ranef.food.wide.tax$kegg.name <- paste0(water.ranef.food.wide.tax$id,"_",water.ranef.food.wide.tax$names)

water.ranef.all1 <- merge(water.ranef.micr.wide.tax,water.ranef.food.wide.tax,by="kegg.name")
any(water.ranef.all1$Pathway.x != water.ranef.all1$Pathway.y)

# Remove perfectly duplicated columns (keep only one copy)
water.ranef.all <- water.ranef.all1 %>%
  select(!matches("\\.y$")) %>%                        # Remove .y columns (from the second df)
  rename_with(~ sub("\\.x$", "", .x), ends_with(".x"))  # Remove .x suffix from first df

water.ranef.all$BvL <- water.ranef.all$Bamboo-water.ranef.all$`Larval food`

#write.csv(water.ranef.all,file="water.fxn.ranef.processed.csv")
```

## Larvae

```{r}
ps.fxn.lar.trim <- readRDS("ps.fxn.lar.trim.rds")

table(larv.mod.ranef$grpvar)

larv.ranef.food <- subset(larv.mod.ranef, grpvar=="variable:Food_type")
larv.ranef.micr <- subset(larv.mod.ranef, grpvar=="variable:Microbe_treatment")
table(larv.ranef.food$grpvar)
table(larv.ranef.micr$grpvar)

larv.ranef.food$id <- str_split_fixed(larv.ranef.food$grp,":",n=2)[,1]
larv.ranef.food$food <- str_split_fixed(larv.ranef.food$grp,":",n=2)[,2]

larv.ranef.micr$id <- str_split_fixed(larv.ranef.micr$grp,":",n=2)[,1]
larv.ranef.micr$microbes <- str_split_fixed(larv.ranef.micr$grp,":",n=2)[,2]

##just the columns for stuff below:
larv.ranef.food.less <- larv.ranef.food %>%
  select(-component,-term,-grp,-condsd)

larv.ranef.micr.less <- larv.ranef.micr %>%
  select(-component,-term,-grp,-condsd)

# Reshape the data frame to have the food types as two columns
larv.ranef.food.wide <- larv.ranef.food.less %>%
  pivot_wider(
    names_from = food,   # Column to use for new column names
    values_from = condval       # Column to fill the new columns
  )

larv.ranef.micr.wide <- larv.ranef.micr.less %>%
  pivot_wider(
    names_from = microbes,   # Column to use for new column names
    values_from = condval       # Column to fill the new columns
  )

##add kegg info
fxn.larv.trim <- as.data.frame(ps.fxn.lar.trim@tax_table)
fxn.larv.trim$id <- fxn.larv.trim$KO

larv.ranef.micr.wide.tax <- merge(larv.ranef.micr.wide,fxn.larv.trim,by="id")
larv.ranef.food.wide.tax <- merge(larv.ranef.food.wide,fxn.larv.trim,by="id")
##more informative kegg ids
larv.ranef.micr.wide.tax$kegg.name <- paste0(larv.ranef.micr.wide.tax$id,"_",larv.ranef.micr.wide.tax$names)
larv.ranef.food.wide.tax$kegg.name <- paste0(larv.ranef.food.wide.tax$id,"_",larv.ranef.food.wide.tax$names)

larv.ranef.all1 <- merge(larv.ranef.micr.wide.tax,larv.ranef.food.wide.tax,by="kegg.name")
any(larv.ranef.all1$Pathway.x != larv.ranef.all1$Pathway.y)

# Remove perfectly duplicated columns (keep only one copy)
larv.ranef.all <- larv.ranef.all1 %>%
  select(!matches("\\.y$")) %>%                        # Remove .y columns (from the second df)
  rename_with(~ sub("\\.x$", "", .x), ends_with(".x"))  # Remove .x suffix from first df

larv.ranef.all$BvL <- larv.ranef.all$Bamboo-larv.ranef.all$`Larval food`

#write.csv(larv.ranef.all,file="larvae.fxn.ranef.processed.csv")
```

# Top 10%

```{r}
water.ranef.all$sBvL <- scale(water.ranef.all$BvL)

Wbamboo <- subset(water.ranef.all,  sBvL > 1.28 & MEM > 0) #top 10%, one-tailed Z score test
dim(Wbamboo)
Wstandard <- subset(water.ranef.all,  sBvL < -1.28 & MMO > 0) #bottom 10%, one-tailed Z score test
dim(Wstandard)

#c(Wbamboo$kegg.name) %in% c(Lbamboo$kegg.name)

##scale BvL column
larv.ranef.all$sBvL <- scale(larv.ranef.all$BvL)

#top.lar <- subset(larv.ranef.all, scale(BvL) > 1.282 | scale(BvL) < -1.282) #top 10%, one-tailed Z score test 
#top.wat <- subset(larv.ranef.all, scale(BvL) > 1.282 | scale(BvL) < -1.282) #top 10%, one-tailed Z score test 
Lbamboo <- subset(larv.ranef.all,  sBvL > 1.28 & MEM > 0) #top 10%, one-tailed Z score test
dim(Lbamboo)
Lstandard <- subset(larv.ranef.all,  sBvL < -1.28 & MMO > 0) #bottom 10%, one-tailed Z score test
dim(Lstandard)
```

# Heat map of top 10% summed by functions

```{r}
library("pheatmap")

topfuns <- read.csv("topfunctiontab.csv",row.names=1)
##get rid of count and sum for the heat map
topfuns2 <- topfuns[row.names(topfuns)!="Count"&row.names(topfuns)!="Total",colnames(topfuns)!="X.1"]

pheatmap(topfuns2,
         scale="row",
         cluster_cols=F)
```

## Nicer heat map

```{r}
colnames(topfuns2) <- c("Larvae - bamboo","Water - bamboo", "Larvae - standard", "Water - standard")

# Draw the heatmap
pheat.topfuns <- pheatmap(topfuns2,
         #color = my_palette,
         scale="row",
         color = colorRampPalette(c("wheat4","white","mediumorchid4"))(100),
         #breaks = breaks,
         #annotation_row = row.source,
         #annotation_colors = ann_colors,
         #scale="column",
         clustering_distance_rows = "maximum",
         border_color = F,
         cluster_cols = F)
```
