---
title: "Mixed models plotting"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Setup

## Libs, directory

```{r}
library("stringr")
library("plyr")
library("dplyr")
library("tidyr")
library("cowplot")
library("ggplot2")
library("cowplot")
library("pheatmap")
library("RColorBrewer")
library(colorspace)
library("phyloseq")
library("pheatmap")

setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/04.mixed_models")
```

## Data

```{r}
#larv.ranef <- read.csv("larv.mod.ranef.csv")
#larv.ranef <- read.csv("larv.mod.ranef.all.csv")
#larv.ranef <- larv.mod.ranef.2ways
larv.ranef <- read.csv("larv.mod.ranef.2ways.csv",row.names=1)

##ps object & taxa table
#ps.lar.trim <- readRDS("ps.exp.lar.trim.rds")
ps.lar.trim <- readRDS("ps.exp.all.lar.trim.rds")
ps.lar.trim #54 taxa, 50 samples
tax.lar.trim <- data.frame(ps.lar.trim@tax_table)
#tax.lar.trim
```

## Checking raw stuff

```{r}
##individual ASVs in all treats
##raw counts
ps.lar.trim.asv <- subset_taxa(ps.lar.trim,id=="ASV0006")
plot_bar(ps.lar.trim.asv)+
  facet_wrap(Food_type~Microbe_treatment,scales="free",ncol=5)

ps.lar.trim.rel = transform_sample_counts(ps.lar.trim, function(x) x / sum(x))
#ps.lar.trim.rel.trt <- subset_samples(ps.lar.trim.rel,Microbe_treatment=="MEM")
ps.lar.trim.rel.asv <- subset_taxa(ps.lar.trim.rel,id=="ASV0085")

plot_bar(ps.lar.trim.rel.asv)+
  facet_wrap(Food_type~Microbe_treatment,ncol=5,scales="free")+
  ggtitle("ASV0085 - Comamonas")
```

### Reformatting ranef table

```{r}
table(larv.ranef$grpvar)

larv.ranef.food <- subset(larv.ranef, grpvar=="variable:Food_type")
larv.ranef.micr <- subset(larv.ranef, grpvar=="variable:Microbe_treatment")
table(larv.ranef.food$grpvar)

# larv.ranef.treat$id <- substr(larv.ranef.treat$grp,1,7)
# larv.ranef.treat$food <- str_sub(larv.ranef.treat$grp,13,18)
# larv.ranef.treat$microbes <- str_sub(larv.ranef.treat$grp,9,11)
# larv.ranef.treat$food_micr <- paste0(larv.ranef.treat$food,"_",larv.ranef.treat$microbes)
#larv.ranef.treat$food_micr <- substr(larv.ranef.treat$grp,9,18)

larv.ranef.food$id <- str_split_fixed(larv.ranef.food$grp,":",n=2)[,1]
larv.ranef.food$food <- str_split_fixed(larv.ranef.food$grp,":",n=2)[,2]

larv.ranef.micr$id <- str_split_fixed(larv.ranef.micr$grp,":",n=2)[,1]
larv.ranef.micr$microbes <- str_split_fixed(larv.ranef.micr$grp,":",n=2)[,2]

##just the columns for stuff below:
larv.ranef.food.less <- larv.ranef.food %>%
  select(-component,-term,-grp,-condsd)

larv.ranef.micr.less <- larv.ranef.micr %>%
  select(-component,-term,-grp,-condsd)

# Reshape the data frame to have the food types as two columns
larv.ranef.food.wide <- larv.ranef.food.less %>%
  pivot_wider(
    names_from = food,   # Column to use for new column names
    values_from = condval       # Column to fill the new columns
  )

larv.ranef.micr.wide <- larv.ranef.micr.less %>%
  pivot_wider(
    names_from = microbes,   # Column to use for new column names
    values_from = condval       # Column to fill the new columns
  )

##add taxa info
larv.ranef.micr.wide.tax <- merge(larv.ranef.micr.wide,tax.lar.trim,by="id")
larv.ranef.food.wide.tax <- merge(larv.ranef.food.wide,tax.lar.trim,by="id")
##more informative ASV ids
larv.ranef.micr.wide.tax$asv.gen <- paste0(larv.ranef.micr.wide.tax$id,"_",larv.ranef.micr.wide.tax$Genus)
larv.ranef.food.wide.tax$asv.gen <- paste0(larv.ranef.food.wide.tax$id,"_",larv.ranef.food.wide.tax$Genus)
```

# Uncorrected conditional modes

```{r}
library("pheatmap")
library("RColorBrewer")

cond.micr.df <- larv.ranef.micr.wide.tax[,c("asv.gen","ECO","MMO","EMO","MEM","ALL")]

#subset(cond.df)

row.names(cond.micr.df) <- cond.micr.df$asv.gen
cond.micr.df2 <- cond.micr.df[,colnames(cond.micr.df)!="asv.gen"]
cond.micr.mat <- as.matrix(cond.micr.df2)

#row.names(otu.rel) == row.names(sam.rel)

max(cond.micr.mat)
min(cond.micr.mat)
# Get symmetric range around 0

abs_max <- max(abs(min_val), abs(max_val))

my_palette <- colorRampPalette(rev(brewer.pal(n=7, name="PiYG")))(10)
breaks <- seq(-abs_max, abs_max, length.out = 11)

# Actual data limits
min_val <- min(cond.micr.mat, na.rm = TRUE)
max_val <- max(cond.micr.mat, na.rm = TRUE)

# Define number of colors on each side of 0 (you can adjust this)
n_neg <- round(100 * abs(min_val) / (abs(min_val) + max_val))  # e.g., ~33
n_pos <- 100 - n_neg                                           # e.g., ~67

# Define color gradients
neg_cols <- colorRampPalette(hcl.colors(7, "Heat 2")[1:4])(n_neg)  # pink to white
pos_cols <- colorRampPalette(hcl.colors(7, "Heat 2")[4:7])(n_pos)  # white to green
my_palette <- c(neg_cols, pos_cols)

# Define breaks so that 0 falls exactly at the transition
breaks <- seq(min_val, max_val, length.out = length(my_palette) + 1)

# Draw the heatmap
pheatmap(cond.micr.mat,
         #color = my_palette,
        color = colorRampPalette(c("blue","white","red"))(100),
         breaks = breaks,
         cluster_cols=F)

library(pheatmap)

# Define your matrix
mat <- cond.micr.mat

# Get data range
min_val <- min(mat, na.rm = TRUE)
max_val <- max(mat, na.rm = TRUE)

# Safety check in case 0 is outside the data range
if (min_val > 0 | max_val < 0) stop("Data does not span 0!")

# Set number of colors (feel free to increase for smoothness)
n_total <- 100

# Calculate number of breaks below and above 0 based on data
n_neg <- round(n_total * abs(min_val) / (abs(min_val) + max_val))
n_pos <- n_total - n_neg

# Create separate color gradients
pink_gradient <- colorRampPalette(c("#FEE0D2", "#FCBBA1", "#FB6A4A", "#EF3B2C", "#CB181D"))(100)

neg_colors <- colorRampPalette(c("white", "black"))(n_neg)
pos_colors <- colorRampPalette(c("black", "#F14C7C"))(n_pos)
full_palette <- c(neg_colors, pos_colors)

# Create matching breaks (length = #colors + 1)
breaks <- c(seq(min_val, 0, length.out = n_neg + 1),
            seq(0, max_val, length.out = n_pos + 1)[-1])  # remove duplicate 0

# Plot!
pheatmap(mat,
         color = full_palette,
         breaks = breaks,
         #scale="row",
         cluster_cols = FALSE,
         border_color=NA)

# Plot!
heatmap <- pheatmap(mat,
         color = colorRampPalette(c("black", "white", "#F14C7C"))(100),
         #breaks = breaksb
         #scale="row",
         cluster_cols = FALSE,
         border_color=NA)

library("colorBlindness")
cvdPlot(heatmap)


ramp <- colorRampPalette(c("#F14C7C", "white", "black"))(100)  # You can also try "Protan" or "Tritan"
cvdPlot()
show_col(ramp)
cvdPlot(ramp)

c(
  "#FF4D92",  # bright fuchsia-pink
  "#FF007F",  # vivid magenta
  "#D5006D",  # strong pink with purple hue
  "#F14C7C",  # rich rose pink
  "#FF3385"   # neon-pink
)

pheatmap(my_matrix,
         color = colorRampPalette(c("#D730A1", "white", "#999999"))(100),
         cluster_cols = FALSE)


(
  "#FEE0D2",  # light pink
  "#FCBBA1",  # peachy pink
  "#FB6A4A",  # coral pink
  "#EF3B2C",  # rose pink
  "#CB181D",  # hot pink
  "#E377C2",  # fuchsia
  "#D4A5A5",  # dusty rose
  "#F781BF"   # magenta-ish
)

# Define your color palette
my_palette = colorRampPalette(brewer.pal(n=9,name="PuRd"))(100)

# Create a ggplot object displaying the color palette
palette_data <- data.frame(
  color = my_palette,
  x = 1:length(my_palette)
)

gg.pal <- ggplot(palette_data, aes(x = x, y=1, fill = color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity() +
  theme_void() +
  labs(title = "Color Palette for cvdPlot")

cvdPlot(gg.pal)

row.source <- data.frame(likely_source=larv.ranef.micr.wide.tax$likely_source, row.names=larv.ranef.micr.wide.tax$asv.gen)

ann_colors = list(likely_source=c(ENV="#84D1B5",LAB="#7D0112",Lab_water="grey80",MOS="#C6A947",`MOS/ENV`="#0085A6"))

pheatmap(cond.micr.mat,
         #scale="row",
         #cluster_cols=F,
        #color = colorRampPalette(c("powderblue", "white","violetred4"))(100),
        annotation_row = row.source,
        annotation_colors = ann_colors,
        border_color = F,
        #color = colorRampPalette(brewer.pal(n=9,name="Purple-Blue"))(100)
        color = colorRampPalette(rev(hcl.colors(n=10,"PinkYl")))(100)
)


# row_annotation <- data.frame(
#   Treatment = c("Control", "Treated", "Treated", "Control"),
#   Timepoint = c("Day1", "Day1", "Day2", "Day2")
# )

# Now pass the annotation to pheatmap
pheatmap(cond.mat.mmo,
         color = colorRampPalette(c("black", "white", "#C6A947"))(100),
         annotation_row = row_annotation
)



```

## Checking origin sources?

```{r}
##MMO origin?
larv.ranef.micr.wide.tax$origin.mmo <- (larv.ranef.micr.wide.tax$MMO/larv.ranef.micr.wide.tax$MEM) - (larv.ranef.micr.wide.tax$EMO/larv.ranef.micr.wide.tax$MEM)

##EMO origin? 
larv.ranef.micr.wide.tax$origin.emo <- (larv.ranef.micr.wide.tax$EMO/larv.ranef.micr.wide.tax$MEM) - (larv.ranef.micr.wide.tax$MMO/larv.ranef.micr.wide.tax$MEM)

ggplot(larv.ranef.micr.wide.tax,aes(x=reorder(asv.gen,origin.mmo),y=origin.mmo,fill=likely_source))+
  geom_bar(stat="identity")+
  theme(axis.text.x=element_text(angle=90,hjust=1))

ps.lar.sourceglom <- tax_glom(ps.lar.trim.rel,"likely_source")

plot_bar(ps.lar.sourceglom,fill="Microbe_treatment",x="likely_source")+
  facet_wrap(~Microbe_treatment,scales="free")
```



# Conditional modes by treatments

## Calculate differences

### Between foods

```{r}
larv.ranef.food.wide.tax$diff.bvl <- larv.ranef.food.wide.tax$Bamboo - larv.ranef.food.wide.tax$`Larval food`

##MMO - MMO BETWEEN FOODS
larv.ranef.wide.tax$diff.bmmo.lmmo <- larv.ranef.wide.tax$Bamboo_MMO - larv.ranef.wide.tax$Larval_MMO
##EMO - EMO BETWEEN FOODS
larv.ranef.wide.tax$diff.bemo.lemo <- larv.ranef.wide.tax$Bamboo_EMO - larv.ranef.wide.tax$Larval_EMO
##MEM - MEM BETWEEN FOODS
larv.ranef.wide.tax$diff.bmem.lmem <- larv.ranef.wide.tax$Bamboo_MEM - larv.ranef.wide.tax$Larval_MEM

##All & eco
##EMO - EMO BETWEEN FOODS
larv.ranef.wide.tax$diff.ball.lall <- larv.ranef.wide.tax$Bamboo_ALL - larv.ranef.wide.tax$Larval_ALL
##MEM - MEM BETWEEN FOODS
larv.ranef.wide.tax$diff.beco.leco <- larv.ranef.wide.tax$Bamboo_ECO - larv.ranef.wide.tax$Larval_ECO

##MMO - EMO BETWEEN FOODS
# larv.ranef.wide.tax$diff.bmmo.lemo <- larv.ranef.wide.tax$Larval_MMO - larv.ranef.wide.tax$Bamboo_EMO
# ##MMO - MEM BETWEEN FOODS
# larv.ranef.wide.tax$diff.bmmo.lmem <- larv.ranef.wide.tax$Larval_MMO - larv.ranef.wide.tax$Bamboo_MEM

##EMO - MMO BETWEEN FOODS
#larv.ranef.wide.tax$diff.lemo.bmmo <- larv.ranef.wide.tax$Larval_EMO - larv.ranef.wide.tax$Bamboo_MMO

##EMO - MEM BETWEEN FOODS
#larv.ranef.wide.tax$diff.lemo.bmem <- larv.ranef.wide.tax$Larval_EMO - larv.ranef.wide.tax$Bamboo_MEM

##MEM - MMO BETWEEN FOODS
#larv.ranef.wide.tax$diff.lmem.bmmo <- larv.ranef.wide.tax$Larval_MEM - larv.ranef.wide.tax$Bamboo_MMO
##MEM - EMO BETWEEN FOODS
#larv.ranef.wide.tax$diff.lmem.bemo <- larv.ranef.wide.tax$Larval_MEM - larv.ranef.wide.tax$Bamboo_EMO

```

## Heat map with everything? 

```{r}
larv.ranef.all <- join(larv.ranef.food.wide.tax,larv.ranef.micr.wide.tax,by="asv.gen")

cond.all.df <- larv.ranef.all[,c("asv.gen","ECO","MMO","EMO","MEM","ALL","Bamboo","Larval food")]

#subset(cond.df)

row.names(cond.all.df) <- cond.all.df$asv.gen
cond.all.df2 <- cond.all.df[,colnames(cond.all.df)!="asv.gen"]
cond.all.mat <- as.matrix(cond.all.df2)

##full thing
pheatmap(cond.all.mat,
         #color = my_palette,
        #color = colorRampPalette(c("blue","white","red"))(100),
         #breaks = breaks,
        scale="row",
         cluster_cols=F)

#row.names(otu.rel) == row.names(sam.rel)

row.source <- data.frame(likely_source=larv.ranef.all$likely_source, row.names=larv.ranef.all$asv.gen)

ann_colors = list(likely_source=c(ENV="#84D1B5",LAB="#7D0112",Lab_water="grey80",MOS="#C6A947",`MOS/ENV`="#0085A6"))

# Actual data limits
min_val <- min(cond.all.mat, na.rm = TRUE)
max_val <- max(cond.all.mat, na.rm = TRUE)

# Define number of colors on each side of 0 (you can adjust this)
n_neg <- round(100 * abs(min_val) / (abs(min_val) + max_val))  # e.g., ~33
n_pos <- 100 - n_neg                                           # e.g., ~67

# Define color gradients
# neg_cols <- colorRampPalette(hcl.colors(10, "RdGy")[1:5])(n_neg)  # pink to white
# pos_cols <- colorRampPalette(hcl.colors(10, "RdGy")[5:10])(n_pos)  # white to green
# my_palette <- c(neg_cols, pos_cols)

neg_colors <- colorRampPalette(c("grey50", "white"))(n_neg)
pos_colors <- colorRampPalette(c("white","orchid4"))(n_pos)
my_palette <- c(neg_colors, pos_colors)

# Define breaks so that 0 falls exactly at the transition
breaks <- seq(min_val, max_val, length.out = length(my_palette) + 1)

# Draw the heatmap
pheatmap(cond.all.mat,
         color = my_palette,
         #color = colorRampPalette(c("blue","white","red"))(100),
         breaks = breaks,
         annotation_row = row.source,
         annotation_colors = ann_colors,
         border_color = F,
         cluster_cols=F)

ramp <- colorRampPalette(c("black","lightsalmon3", "white", "slategray"))(10)  # You can also try "Protan" or "Tritan"
palette_data <- data.frame(
  color = ramp,
  x = 1:length(ramp)
)

gg.pal <- ggplot(palette_data, aes(x = x, y=1, fill = color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity() +
  theme_void() +
  labs(title = "Color Palette for cvdPlot")

cvdPlot(gg.pal)

library(pheatmap)

# Define your matrix
mat <- cond.micr.mat

# Get data range
min_val <- min(mat, na.rm = TRUE)
max_val <- max(mat, na.rm = TRUE)

# Set number of colors (feel free to increase for smoothness)
n_total <- 100

# Calculate number of breaks below and above 0 based on data
n_neg <- round(n_total * abs(min_val) / (abs(min_val) + max_val))
n_pos <- n_total - n_neg

# Create separate color gradients
pink_gradient <- colorRampPalette(c("#FEE0D2", "#FCBBA1", "#FB6A4A", "#EF3B2C", "#CB181D"))(100)

neg_colors <- colorRampPalette(c("white", "black"))(n_neg)
pos_colors <- colorRampPalette(c("black", "#F14C7C"))(n_pos)
full_palette <- c(neg_colors, pos_colors)

# Create matching breaks (length = #colors + 1)
breaks <- c(seq(min_val, 0, length.out = n_neg + 1),
            seq(0, max_val, length.out = n_pos + 1)[-1])  # remove duplicate 0

# Plot!
pheatmap(mat,
         color = full_palette,
         breaks = breaks,
         #scale="row",
         cluster_cols = FALSE,
         border_color=NA)

# Plot!
heatmap <- pheatmap(mat,
         color = colorRampPalette(c("black", "white", "#F14C7C"))(100),
         #breaks = breaksb
         #scale="row",
         cluster_cols = FALSE,
         border_color=NA)

library("colorBlindness")
cvdPlot(heatmap)


ramp <- colorRampPalette(c("#F14C7C", "white", "black"))(100)  # You can also try "Protan" or "Tritan"
cvdPlot()
show_col(ramp)
cvdPlot(ramp)

c(
  "#FF4D92",  # bright fuchsia-pink
  "#FF007F",  # vivid magenta
  "#D5006D",  # strong pink with purple hue
  "#F14C7C",  # rich rose pink
  "#FF3385"   # neon-pink
)

pheatmap(my_matrix,
         color = colorRampPalette(c("#D730A1", "white", "#999999"))(100),
         cluster_cols = FALSE)


(
  "#FEE0D2",  # light pink
  "#FCBBA1",  # peachy pink
  "#FB6A4A",  # coral pink
  "#EF3B2C",  # rose pink
  "#CB181D",  # hot pink
  "#E377C2",  # fuchsia
  "#D4A5A5",  # dusty rose
  "#F781BF"   # magenta-ish
)

# Define your color palette
my_palette = colorRampPalette(brewer.pal(n=9,name="PuRd"))(100)

# Create a ggplot object displaying the color palette
palette_data <- data.frame(
  color = my_palette,
  x = 1:length(my_palette)
)

gg.pal <- ggplot(palette_data, aes(x = x, y=1, fill = color)) +
  geom_bar(stat = "identity") +
  scale_fill_identity() +
  theme_void() +
  labs(title = "Color Palette for cvdPlot")

cvdPlot(gg.pal)

row.source <- data.frame(likely_source=larv.ranef.micr.wide.tax$likely_source, row.names=larv.ranef.micr.wide.tax$asv.gen)

ann_colors = list(likely_source=c(ENV="#84D1B5",LAB="#7D0112",Lab_water="grey80",MOS="#C6A947",`MOS/ENV`="#0085A6"))

pheatmap(cond.micr.mat,
         #scale="row",
         #cluster_cols=F,
        #color = colorRampPalette(c("powderblue", "white","violetred4"))(100),
        annotation_row = row.source,
        annotation_colors = ann_colors,
        border_color = F,
        #color = colorRampPalette(brewer.pal(n=9,name="Purple-Blue"))(100)
        color = colorRampPalette(rev(hcl.colors(n=10,"PinkYl")))(100)
)


# row_annotation <- data.frame(
#   Treatment = c("Control", "Treated", "Treated", "Control"),
#   Timepoint = c("Day1", "Day1", "Day2", "Day2")
# )

# Now pass the annotation to pheatmap
pheatmap(cond.mat.mmo,
         color = colorRampPalette(c("black", "white", "#C6A947"))(100),
         annotation_row = row_annotation
)


```

### Within foods

```{r}
##MMO - EMO WITHIN BAMBOO
#larv.ranef.wide.tax$diff.bmmo.bemo <- larv.ranef.wide.tax$Bamboo_MMO - larv.ranef.wide.tax$Bamboo_EMO
##MMO - MEM WITHIN BAMBOO
larv.ranef.wide.tax$diff.bmmo.bmem <- larv.ranef.wide.tax$Bamboo_MMO - larv.ranef.wide.tax$Bamboo_MEM
##EMO - MEM WITHIN BAMBOO
larv.ranef.wide.tax$diff.bemo.bmem <- larv.ranef.wide.tax$Bamboo_EMO - larv.ranef.wide.tax$Bamboo_MEM

##MMO - EMO WITHIN LARVAL FOOD
#larv.ranef.wide.tax$diff.lmmo.lemo <- larv.ranef.wide.tax$Larval_MMO - larv.ranef.wide.tax$Larval_EMO
##MMO - MEM WITHIN LARVAL FOOD
larv.ranef.wide.tax$diff.lmmo.lmem <- larv.ranef.wide.tax$Larval_MMO - larv.ranef.wide.tax$Larval_MEM
##EMO - MEM WITHIN LARVAL FOOD
larv.ranef.wide.tax$diff.lemo.lmem <- larv.ranef.wide.tax$Larval_EMO - larv.ranef.wide.tax$Larval_MEM

```

## Saving

```{r}
#write.csv(larv.ranef.wide.tax,file="larv.ranef.wide.tax.all.csv")
#write.csv(larv.ranef.wide.tax,file="larv.ranef.wide.tax.csv")
```

## Checking one out as an example

```{r}
larv.ranef.diffs <- larv.ranef.wide.tax %>%
  select(id, starts_with("diff.")) %>%
  pivot_longer(
    cols = starts_with("diff."),
    names_to = "comparison",
    values_to = "difference"
  )

#larv.ranef.diffs.asv1 <- subset(larv.ranef.diffs,id=="ASV0001")

#ggplot(larv.ranef.diffs.asv1,aes(x=comparison,y=difference))+
#  geom_bar(stat="identity")+
#  theme(axis.text.x=element_text(angle=45,hjust=1))

##alternatively
ggplot(larv.ranef.wide.tax,aes(x=reorder(asv.gen,diff.bemo.lemo),y=diff.bemo.lemo))+
  geom_bar(stat="identity")+
  theme(axis.text.x=element_text(angle=45,hjust=1))+
  ggtitle("EMO - bamboo minus larval food")

##individual ASVs within treatments
ps.lar.trim.rel = transform_sample_counts(ps.lar.trim, function(x) x / sum(x))
ps.lar.trim.rel.trt <- subset_samples(ps.lar.trim.rel,Microbe_treatment=="MEM")
ps.lar.trim.rel.trt.asv <- subset_taxa(ps.lar.trim.rel.trt,id=="ASV0027")

plot_bar(ps.lar.trim.rel.trt.asv)+
  facet_wrap(Food_type~Microbe_treatment)
```

## Plotting{.tabset}

```{r}
larv.ranef.wide.tax <- read.csv("larv.ranef.wide.tax.csv")
```

### Just food

```{r}
ggplot(larv.ranef.food.wide.tax,aes(x=reorder(asv.gen,diff.bvl),y=diff.bvl,fill=likely_source))+
  #geom_bar(stat="identity",position=position_dodge(1))+
  geom_bar(stat="identity",color="black")+
  scale_fill_manual(values=c("#84D1B5","#7D0112","grey80","#C6A947","#0085A6"))+
  theme_cowplot()+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  ggtitle("Bamboo - larval food")
```

### MMO as the reference level

```{r}
larv.ranef.micr.wide.tax$diff.mmo.emo <- larv.ranef.micr.wide.tax$MMO - larv.ranef.micr.wide.tax$EMO
larv.ranef.micr.wide.tax$diff.mmo.mem <- larv.ranef.micr.wide.tax$MMO - larv.ranef.micr.wide.tax$MEM
larv.ranef.micr.wide.tax$diff.mmo.all <- larv.ranef.micr.wide.tax$MMO - larv.ranef.micr.wide.tax$ALL
larv.ranef.micr.wide.tax$diff.mmo.eco <- larv.ranef.micr.wide.tax$MMO - larv.ranef.micr.wide.tax$ECO
```

#### Plot

```{r}
cond.df.mmo <- larv.ranef.micr.wide.tax[,c("asv.gen","diff.mmo.emo","diff.mmo.mem","diff.mmo.all","diff.mmo.eco")]

#subset(cond.df)

row.names(cond.df.mmo) <- cond.df.mmo$asv.gen
cond.df.mmo2 <- cond.df.mmo[,colnames(cond.df.mmo)!="asv.gen"]
cond.mat.mmo <- as.matrix(cond.df.mmo2)

#row.names(otu.rel) == row.names(sam.rel)

# Suppose cond.mat.mmo has samples (rows) and features (columns)
# and your annotation looks like this:
row.source <- data.frame(likely_source=larv.ranef.micr.wide.tax$likely_source, row.names=larv.ranef.micr.wide.tax$asv.gen)
# row_annotation <- data.frame(
#   Treatment = c("Control", "Treated", "Treated", "Control"),
#   Timepoint = c("Day1", "Day1", "Day2", "Day2")
# )

# Now pass the annotation to pheatmap
pheatmap(cond.mat.mmo,
         color = colorRampPalette(c("black", "white", "#C6A947"))(100),
         annotation_row = row_annotation
)

ann_colors = list(
    likely_source=c(ENV="#84D1B5",LAB="#7D0112",Lab_water="grey80",MOS="#C6A947",`MOS/ENV`="#0085A6")
)

pheatmap(cond.mat.mmo,
        scale="row",
         #cluster_cols=F,
        color = colorRampPalette(c("black","white","#C6A947"))(100),
        annotation_row = row.source,
        annotation_colors = ann_colors
        #color = colorRampPalette(rev(brewer.pal(n=7,name="PiYG")))(100)
        #color = colorRampPalette(rev(brewer.pal(n=7,name="PiYG")))(100)
)
#("#84D1B5","#7D0112","grey80","#C6A947","#0085A6"))+

pheatmap(cond.mat,
         scale="row",
         #cluster_cols=F,
        #color = colorRampPalette(c("blue", "white", "red"))(100),
        color = colorRampPalette(rev(brewer.pal(n=7,name="PiYG")))(100)
)
```

### Mosq. microbes

```{r}
ggplot(larv.ranef.wide.tax,aes(x=reorder(asv.gen,diff.bmmo.lmmo),y=diff.bmmo.lmmo))+
  #geom_bar(stat="identity",position=position_dodge(1))+
  geom_bar(stat="identity")+
  #scale_fill_manual(values=c("#C6A947","#0085A6","#84D1B5"))+
  theme_cowplot()+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  ggtitle("MMO - bamboo minus larval food")

```

### Env. microbes

```{r}
ggplot(larv.ranef.wide.tax,aes(x=reorder(asv.gen,diff.bemo.lemo),y=diff.bemo.lemo))+
  #geom_bar(stat="identity",position=position_dodge(1))+
  geom_bar(stat="identity")+
  #scale_fill_manual(values=c("#C6A947","#0085A6","#84D1B5"))+
  theme_cowplot()+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  ggtitle("EMO - bamboo minus larval food")

```

### Mosq + env microbes

```{r}
ggplot(larv.ranef.wide.tax,aes(x=reorder(asv.gen,diff.bmem.lmem),y=diff.bmem.lmem))+
  #geom_bar(stat="identity",position=position_dodge(1))+
  geom_bar(stat="identity")+
  #scale_fill_manual(values=c("#C6A947","#0085A6","#84D1B5"))+
  theme_cowplot()+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  ggtitle("MEM - bamboo minus larval food")
```

## Heat map

```{r}
library("pheatmap")
library("RColorBrewer")

#diff.df <- larv.ranef.wide.tax[,c("asv.gen","diff.beco.leco","diff.bemo.lemo","diff.bmmo.lmmo","diff.bmem.lmem","diff.ball.lall")]
diff.df <- larv.ranef.wide.tax[,c("asv.gen","diff.bemo.lemo","diff.bmmo.lmmo","diff.bmem.lmem")]

row.names(diff.df) <- diff.df$asv.gen
diff.df2 <- diff.df[,2:4]
#diff.df2 <- diff.df[,2:6]
diff.mat <- as.matrix(diff.df2)

#row.names(otu.rel) == row.names(sam.rel)

pheatmap(diff.mat,
        color = colorRampPalette(c("#8C510A", "white", "darkolivegreen4"))(100),
        scale="row",
        cluster_rows=T,
        cluster_cols=T,
        #color = colorRampPalette(brewer.pal(n=7,name="BrBG"))(100)
        #color = colorRampPalette(brewer.pal(n=7,name="PiYG"))(100)
        #color = colorRampPalette(rev(brewer.pal(n=7,name="PiYG")))(100)
)

"#8C510A"
```

## Plotting - all treatments{.tabset}

```{r}
larv.ranef.wide.tax <- read.csv("larv.ranef.wide.tax.all.csv")
```

### Mosq. microbes

```{r}
ggplot(larv.ranef.wide.tax,aes(x=reorder(asv.gen,diff.bmmo.lmmo),y=diff.bmmo.lmmo))+
  #geom_bar(stat="identity",position=position_dodge(1))+
  geom_bar(stat="identity")+
  #scale_fill_manual(values=c("#C6A947","#0085A6","#84D1B5"))+
  theme_cowplot()+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  ggtitle("MMO - bamboo minus larval food")

```

### Env. microbes

```{r}
ggplot(larv.ranef.wide.tax,aes(x=reorder(asv.gen,diff.bemo.lemo),y=diff.bemo.lemo))+
  #geom_bar(stat="identity",position=position_dodge(1))+
  geom_bar(stat="identity")+
  #scale_fill_manual(values=c("#C6A947","#0085A6","#84D1B5"))+
  theme_cowplot()+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  ggtitle("EMO - bamboo minus larval food")

```

### Mosq + env microbes

```{r}
ggplot(larv.ranef.wide.tax,aes(x=reorder(asv.gen,diff.bmem.lmem),y=diff.bmem.lmem))+
  #geom_bar(stat="identity",position=position_dodge(1))+
  geom_bar(stat="identity")+
  #scale_fill_manual(values=c("#C6A947","#0085A6","#84D1B5"))+
  theme_cowplot()+
  theme(axis.text.x=element_text(angle=90,hjust=1))+
  ggtitle("MEM - bamboo minus larval food")
```

## Heat map

```{r}
library("pheatmap")
library("RColorBrewer")

diff.df <- larv.ranef.wide.tax[,c("asv.gen","diff.beco.leco","diff.bemo.lemo","diff.bmmo.lmmo","diff.bmem.lmem","diff.ball.lall")]
#diff.df <- larv.ranef.wide.tax[,c("asv.gen","diff.bemo.lemo","diff.bmmo.lmmo","diff.bmem.lmem")]

row.names(diff.df) <- diff.df$asv.gen
#diff.df2 <- diff.df[,2:4]
diff.df2 <- diff.df[,2:6]
diff.mat <- as.matrix(diff.df2)

#row.names(otu.rel) == row.names(sam.rel)

pheatmap(diff.mat,
        color = colorRampPalette(c("#8C510A", "white", "darkolivegreen4"))(100),
        scale="row",
        cluster_rows=T,
        cluster_cols=T,
        #color = colorRampPalette(brewer.pal(n=7,name="BrBG"))(100)
        #color = colorRampPalette(brewer.pal(n=7,name="PiYG"))(100)
        #color = colorRampPalette(rev(brewer.pal(n=7,name="PiYG")))(100)
)

"#8C510A"
```

# Stopped updating script here

# ggpredict

```{r, eval=F}
#install.packages("ggeffects")
library("ggeffects")

setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/04.mixed_models")

larv.mod.nb1 <- readRDS("larv.mod.nb1.rds")

#ps.trim.wtr <- readRDS("ps.exp.wtr.trim.rds")
#ps.trim.wtr
ps.trim.lar <- readRDS("ps.exp.lar.trim.rds")
ps.trim.lar
```

```{r, eval=F}
pred.lar.nb1 <- ggpredict(larv.mod.nb1, terms=c("variable", "Food_type","Microbe_treatment"), type="re")

plot(pred.lar.nb1)+
  theme(axis.text.x=element_text(angle=90))+
  theme_classic()

# pred.lar.nb2 <- ggpredict(larv.mod.nb2, terms=c("variable", "Food_type","Microbe_treatment"), type="re")
# plot(pred.lar.nb2)

tax.trim.lar <- data.frame(ps.trim.lar@tax_table)

##make same column names
pred.lar.nb1$id <- pred.lar.nb1$x
pred.lar.tax <- merge(tax.trim.lar,pred.lar.nb1,by="id")

ggplot(pred.lar.tax,aes(x=group,y=predicted,color=facet))+
  geom_point()+
  facet_wrap(~facet)

pred.lar.tax.mmo <- subset(pred.lar.tax,facet=="MMO")

ggplot(pred.lar.tax.mmo,aes(x=group,y=predicted,fill=id))+
  #geom_bar(stat="identity",position=position_dodge(0.5))+
  geom_bar(stat="identity")+
  facet_wrap(~Genus,scales="free")+
  theme(legend.position="none")

pred.lar.tax.bamb <- subset(pred.lar.tax,group=="Bamboo")

ggplot(pred.lar.tax.bamb,aes(x=facet,y=predicted,fill=id))+
  geom_bar(stat="identity")+
  facet_wrap(~Family,scales="free")+
  theme(legend.position="none")
```

Trying some different plots

```{r, eval=F}
library("cowplot")

pred.lar.tax.comp <- subset(pred.lar.tax,facet!="ECO"&facet!="ALL")

pred.lar.tax.comp.bam <- subset(pred.lar.tax.comp,group=="Bamboo")

ggplot(pred.lar.tax.comp.bam,aes(x=facet,y=predicted,fill=facet))+
  facet_wrap(~id,scales="free")+
  geom_bar(stat="identity")+
  theme_cowplot()

##heatmap
library("pheatmap")
library("RColorBrewer")

library("tidyr")

#pred.lar.tax.comp$asv_food <- paste0(pred.lar.tax.comp$id,"_",pred.lar.tax.comp$group)

colnames(pred.lar.tax.comp)

##remove not needed columns
pred.lar.min <- pred.lar.tax.comp %>%
  select(predicted,facet,group,id)

##bamboo
pred.lar.min.bam <- subset(pred.lar.min,group=="Bamboo")

pred.lar.min.bamw <- pred.lar.min.bam %>%
  pivot_wider(names_from=id,values_from=predicted) %>%
  as.data.frame()

row.names(pred.lar.min.bamw) <- pred.lar.min.bamw$facet

pred.lar.min.bamw2 <- pred.lar.min.bamw[,3:ncol(pred.lar.min.bamw)]

pheatmap(as.matrix(pred.lar.min.bamw2),
        color = colorRampPalette(c("white", "orange", "red"))(100),
        #color = colorRampPalette(rev(brewer.pal(n=7,name="PiYG")))(100),
        #scale="column",
        cluster_cols = F,
        cluster_rows=F
)

##larval food
pred.lar.min.lar <- subset(pred.lar.min,group=="Larval food")

pred.lar.min.larw <- pred.lar.min.lar %>%
  pivot_wider(names_from=id,values_from=predicted) %>%
  as.data.frame()

row.names(pred.lar.min.larw) <- pred.lar.min.larw$facet

pred.lar.min.larw2 <- pred.lar.min.larw[,3:ncol(pred.lar.min.larw)]

pheatmap(as.matrix(pred.lar.min.larw2),
        color = colorRampPalette(c("white", "orange", "red"))(100),
        #color = colorRampPalette(rev(brewer.pal(n=7,name="PiYG")))(100),
        #scale="column",
        cluster_cols = F,
        cluster_rows=F
)

library("ggrepel")

##correlating what happened in bamboo & larval food
pred.lar.min$group <- gsub("Larval food","Larval",pred.lar.min$group)

pred.lar.min.cor <- pred.lar.min %>%
  pivot_wider(names_from=group,values_from=predicted)

pred.lar.min.cor$facet <- factor(pred.lar.min.cor$facet,levels=c("MMO","EMO","MEM"))

##full plot
ggplot(pred.lar.min.cor,aes(x=Bamboo,y=Larval,color=facet,shape=facet,label=id))+
  geom_point(size=2)+
  geom_abline(intercept=0,slope=1,linetype="dotted")+
  geom_text_repel(color="black")+
  theme_cowplot()+
  ylab("Standard diet")+
  xlab("Bamboo leaves")+
  scale_color_manual(values=c("#C6A947","#84D1B5","#0085A6"))

##zooming into the low section
ggplot(pred.lar.min.cor,aes(x=Bamboo,y=Larval,color=facet,shape=facet,label=id))+
  geom_point()+
  geom_abline(intercept=0,slope=1,linetype="dotted")+
  geom_abline(intercept=0,slope=0.5,linetype="dotted")+
  geom_abline(intercept=0,slope=2,linetype="dotted")+
  xlim(0,100)+
  ylim(0,100)+
  geom_text_repel(color="black")+
  theme_cowplot()+
  ylab("Standard diet")+
  xlab("Bamboo leaves")+
  scale_color_manual(values=c("#C6A947","#84D1B5","#0085A6"))

mean(sample_sums(ps.lar.trim))
#47762.4
1000/47762.4

##max 2k - larval food
ggplot(pred.lar.min.cor,aes(x=Bamboo,y=Larval,color=facet,shape=facet,label=id))+
  geom_point()+
  geom_abline(intercept=0,slope=1,linetype="dotted")+
  geom_abline(intercept=0,slope=0.5,linetype="dotted")+
  geom_abline(intercept=0,slope=2,linetype="dotted")+
  #xlim(0,4000)+
  ylim(1000,30000)+
  geom_text_repel(color="black")+
  theme_cowplot()+
  ylab("Standard diet")+
  xlab("Bamboo leaves")+
  scale_color_manual(values=c("#C6A947","#84D1B5","#0085A6"))

##max 2k - bamboo
ggplot(pred.lar.min.cor,aes(x=Bamboo,y=Larval,color=facet,shape=facet,label=id))+
  geom_point()+
  geom_abline(intercept=0,slope=1,linetype="dotted")+
  geom_abline(intercept=0,slope=0.5,linetype="dotted")+
  geom_abline(intercept=0,slope=2,linetype="dotted")+
  xlim(1000,30000)+
  #ylim(1000,30000)+
  geom_text_repel(color="black")+
  theme_cowplot()+
  ylab("Standard diet")+
  xlab("Bamboo leaves")+
  scale_color_manual(values=c("#C6A947","#84D1B5","#0085A6"))
```

Checking what the actual data look like

```{r, eval=F}
ps.trim.lar.comp <- subset_samples(ps.lar.trim,Microbe_treatment!="ALL"&Microbe_treatment!="ECO")
ps.trim.lar.comp

ps.lar.trim.rel = transform_sample_counts(ps.trim.lar.comp, function(x) x / sum(x))
ps.lar.trim.rel

ps.asv1 <- subset_taxa(ps.trim.lar.comp,id=="ASV0001")
ps.asv1@otu_table
plot_bar(ps.asv1,x="Sample")+
  facet_wrap(Microbe_treatment~Food_type)
plot_bar(ps.asv1)+
  facet_wrap(Microbe_treatment~Food_type,scales="free")

ps.asv3 <- subset_taxa(ps.lar.trim.rel,id=="ASV0003")
ps.asv3
plot_bar(ps.asv3,x="Microbe_treatment")+
  facet_wrap(~Food_type)
plot_bar(ps.asv3)+
  facet_wrap(Microbe_treatment~Food_type,scales="free")

ps.asv4 <- subset_taxa(ps.lar.trim.rel,id=="ASV0004")
ps.asv4
plot_bar(ps.asv4,x="Microbe_treatment")+
  facet_wrap(~Food_type)

ps.asv6 <- subset_taxa(ps.lar.trim.rel,id=="ASV0006")
ps.asv6
plot_bar(ps.asv6,x="Microbe_treatment")+
  facet_wrap(~Food_type)

ps.asv7 <- subset_taxa(ps.lar.trim.rel,id=="ASV0007")
ps.asv7
plot_bar(ps.asv7,x="Microbe_treatment")+
  facet_wrap(~Food_type)

ps.asv19 <- subset_taxa(ps.trim.lar.comp,id=="ASV0019")
ps.asv19
plot_bar(ps.asv19,x="Microbe_treatment")+
  facet_wrap(~Food_type)
plot_bar(ps.asv19)+
  facet_wrap(Microbe_treatment~Food_type,scales="free")

ps.asv33 <- subset_taxa(ps.trim.lar.comp,id=="ASV0033")
ps.asv33
plot_bar(ps.asv33,x="Microbe_treatment")+
  facet_wrap(~Food_type)
plot_bar(ps.asv33)+
  facet_wrap(Microbe_treatment~Food_type,scales="free")

ps.asv54 <- subset_taxa(ps.lar.trim.rel,id=="ASV0054")
ps.asv54
plot_bar(ps.asv54,x="Microbe_treatment")+
  facet_wrap(~Food_type)
plot_bar(ps.asv54)+
  facet_wrap(Microbe_treatment~Food_type)

larv.ranef.wide.tax[6,]


         
#         )+
#  facet_wrap(Food_type~Microbe_treatment)
```


