---
title: "Bamboo mesocosms - physiology results"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
editor_options: 
  chunk_output_type: console
---

Notes: almost everything is "done" in the code here - except moving the library packages to where they're supposed to be and maybe cleaning up a few places & adding some notes

# Setup

## Libraries

```{r}
library("dplyr")
library("ggplot2")
library("cowplot")

# #install.packages("lubridate")
# library("lubridate")
# #install.packages("Rmisc")
# library("Rmisc")
# #install.packages("coxme")
# library("coxme")
# library("car")
# #install.packages("bbmle")
# library("bbmle")
# #install.packages("emmeans")
# library("emmeans")
# #install.packages("DHARMa")
# library("DHARMa")
# #install.packages("multcomp")
# library("multcomp")
# #install.packages("multcompView")
# library("multcompView")
# library("glmmTMB")
```

## Read in data

```{r}
setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/05.phys_analysis")

pupae.wts <- read.csv("../bamboomesos_pupaeweights.csv")
```

# Biomass by date

```{r}
##doesn't make sense to compare with the AHK & None_larvae microbe treatments since they only had one mesocosm replicate each, but leaving in for now for interest
#pupae.wts.noahk <- subset(pupae.wts,Microbe_treatment!="AHK")
#pupae.wts.exp <- subset(pupae.wts.noahk,Microbe_treatment!="None_larvae")

#pupae.wts$food.microbe <- paste0(pupae.wts$Food_type,"_",pupae.wts$Microbe_treatment)

pupae.wts$Dry_weightAVG[is.na(pupae.wts$Dry_weightAVG)] <- 0

pupae.wts.sum <- pupae.wts %>%
  dplyr::group_by(Food_type,Microbe_treatment,Exp_day) %>%
  dplyr::summarize(biomass_sum=sum(Dry_weightAVG)) %>%
  dplyr::mutate(biomass_cumsum=cumsum(biomass_sum))

pupae.wts.sum2 <- pupae.wts.sum %>%
  as.data.frame() %>%
  add_row(Food_type="Bamboo",Microbe_treatment="ECO",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="EMO",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="MMO",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="MEM",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="ALL",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="AHK",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="None_larvae",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Larval food",Microbe_treatment="ECO",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Larval food",Microbe_treatment="EMO",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Larval food",Microbe_treatment="MMO",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Larval food",Microbe_treatment="MEM",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Larval food",Microbe_treatment="ALL",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Larval food",Microbe_treatment="AHK",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Larval food",Microbe_treatment="None_larvae",Exp_day=0,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="ECO",Exp_day=40,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="EMO",Exp_day=40,biomass_sum=0,biomass_cumsum=8.963) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="MMO",Exp_day=40,biomass_sum=0,biomass_cumsum=7.355) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="MEM",Exp_day=40,biomass_sum=0,biomass_cumsum=9.671) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="ALL",Exp_day=40,biomass_sum=0,biomass_cumsum=11.406) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="AHK",Exp_day=40,biomass_sum=0,biomass_cumsum=0) %>%
  add_row(Food_type="Bamboo",Microbe_treatment="None_larvae",Exp_day=40,biomass_sum=0,biomass_cumsum=0.197) %>%
  add_row(Food_type="Larval food",Microbe_treatment="ECO",Exp_day=40,biomass_sum=0,biomass_cumsum=31.091) %>%
  add_row(Food_type="Larval food",Microbe_treatment="EMO",Exp_day=40,biomass_sum=0,biomass_cumsum=38.516) %>%
  add_row(Food_type="Larval food",Microbe_treatment="MMO",Exp_day=40,biomass_sum=0,biomass_cumsum=65.280) %>%
  add_row(Food_type="Larval food",Microbe_treatment="MEM",Exp_day=40,biomass_sum=0,biomass_cumsum=17.002) %>%
  add_row(Food_type="Larval food",Microbe_treatment="ALL",Exp_day=40,biomass_sum=0,biomass_cumsum=15.672) %>%
  add_row(Food_type="Larval food",Microbe_treatment="AHK",Exp_day=40,biomass_sum=0,biomass_cumsum=15.548) %>%
  add_row(Food_type="Larval food",Microbe_treatment="None_larvae",Exp_day=40,biomass_sum=0,biomass_cumsum=7.414)

pupae.wts.sum3 <- subset(pupae.wts.sum2,Microbe_treatment!="None_larvae"&Microbe_treatment!="AHK")

pupae.wts.sum3$biomass_cumsum_by5 <- (pupae.wts.sum3$biomass_cumsum)/5

#new.colors <- c("gray60","#D83706","#3B9FFD","#4145AA","black")

pupae.wts.sum3$Microbe_treatment <- factor(pupae.wts.sum3$Microbe_treatment,levels=c("ECO","MMO","EMO","MEM","ALL"))

##Matt versions
#install.packages("extrafont")
# library("extrafont")
# #install.packages("extrafontdb")
# library("extrafontdb")
# font_import()

new.colors <- c("#7D0112","#C6A947","#84D1B5","#0085A6","#1F28A2")

##average by 5
ggplot(pupae.wts.sum3,aes(x=Exp_day,y=biomass_cumsum_by5,color=Microbe_treatment,linetype=Food_type))+
  #geom_vline(xintercept=5,linetype="dotted",alpha=0.5)+
  #geom_vline(xintercept=8,linetype="dotted",alpha=0.5)+
  geom_step(lwd=0.9)+
  #scale_linetype_manual(name="Diet",labels=c("Bamboo","Lab food"),values=c("solid","twodash"))+
  #xlim(0,39)+
  #theme_cowplot()+
  theme_linedraw()+
  xlab("Experimental day")+
  ylab("Cumulative pupal biomass (mg) ")+
  #scale_y_continuous(labels=c("0 mg","5 mg","10 mg","15 mg"))+
  scale_color_manual(values=new.colors)+
  #scale_color_manual(name="Microbes",values=new.colors,labels=c("Marginal microbiome (MM)","Mosquito symbionts only (MS)","Environmental microbes only (EM)","MS+EM","MS+EM+MM"))+
  #theme(text=element_text(family="Calibri"),legend.key.width = unit(1.85, "line"))+
  scale_linetype_manual(values=c("solid","longdash"),name="Food type",labels=c("Bamboo detritus","Lab food"))
  #scale_color_manual(values=c("#7DBA54","#992559","#60eede"),name="Infusion")

#ggsave("pupae.weights.sum.png",width=5,height=3)

##average by 5
pupae.wts.sum3.bamboo <- subset(pupae.wts.sum3,Food_type=="Bamboo")
gg.bamboo <- ggplot(pupae.wts.sum3.bamboo,aes(x=Exp_day,y=biomass_cumsum_by5,color=Microbe_treatment))+
  #geom_vline(xintercept=5,linetype="dotted",alpha=0.5)+
  #geom_vline(xintercept=8,linetype="dotted",alpha=0.5)+
  geom_step(lwd=0.9)+
  #scale_linetype_manual(name="Diet",labels=c("Bamboo","Lab food"),values=c("solid","twodash"))+
  #xlim(0,39)+
  #theme_cowplot()+
  theme_cowplot()+
  xlab("Experimental day")+
  ylab("Cumulative pupal biomass (mg)")+
  scale_y_continuous(breaks=c(0,1,2),labels=c("0","1","2"))+
  #scale_color_manual(values=new.colors)+
  scale_color_manual(name="Microorganisms",values=new.colors,labels=c("LAB","MOS","ENV","MOS+ENV","MOS+ENV+LAB"))+
  #theme(legend.key.width = unit(1.85, "line"))+
  coord_cartesian(xlim=c(7,40))+
  #geom_hline(yintercept=0,lwd=0.9,color="gray60")+
  #scale_linetype_manual(values=c("solid","longdash"),name="Food type",labels=c("Bamboo detritus","Lab food"))+
  #theme(axis.title.y=element_blank())+
  ylab("")+
  ggtitle("Bamboo diet")
  #scale_color_manual(values=c("#7DBA54","#992559","#60eede"),name="Infusion")

gg.bamboo

#ggsave("pupae.weights.sum_bamboo.png",width=5,height=3)

##average by 5
pupae.wts.sum3.labfood <- subset(pupae.wts.sum3,Food_type=="Larval food")
gg.labfood <- ggplot(pupae.wts.sum3.labfood,aes(x=Exp_day,y=biomass_cumsum_by5,color=Microbe_treatment))+
  #geom_vline(xintercept=5,linetype="dotted",alpha=0.5)+
  #geom_vline(xintercept=8,linetype="dotted",alpha=0.5)+
  geom_step(lwd=0.9)+
  #scale_linetype_manual(name="Diet",labels=c("Bamboo","Lab food"),values=c("solid","twodash"))+
  #xlim(0,39)+
  theme_cowplot()+
  #theme_linedraw()+
  xlab("Experimental day")+
  ylab("Sum of pupal biomass (mg)  ")+
  ggtitle("Standard diet")+
  #scale_y_continuous(labels=c("0 mg","5 mg","10 mg","15 mg"))+
  #scale_color_manual(values=new.colors)+
  scale_color_manual(name="Microorganisms",values=new.colors,labels=c("Lab strains (LAB)","Mosquito (MOS)","Environmental (ENV)","MOS+ENV","MOS+ENV+LAB"))+
  #theme(legend.key.width = unit(1.85, "line"))+
  #theme(axis.title.x=element_blank())+
  coord_cartesian(xlim=c(7,40))
  #geom_hline(yintercept=0,lwd=0.9,color="gray60")+
  #scale_linetype_manual(values=c("solid","longdash"),name="Food type",labels=c("Bamboo detritus","Lab food"))
  #scale_color_manual(values=c("#7DBA54","#992559","#60eede"),name="Infusion")
gg.labfood

#ggsave("pupae.weights.sum_labfood.png",width=5,height=3)

library("ggpubr")

ggarrange(gg.labfood,gg.bamboo,common.legend=T,legend="right",labels=c("(b)","(c)"))
#ggsave("facet.diet.pdf",width=8,height=3)

## survival plot stuff I didn't end up using
# install.packages(c("survival", "lubridate", "ggsurvfit", "gtsummary", "tidycmprsk"))
# remotes::install_github("zabore/condsurv")
# remotes::install_github("zabore/ezfun")
# library(gtsummary)
# library(tidycmprsk)
# library(condsurv)

# library(survival)
# #install.packages("ggsurvfit")
# library(ggsurvfit)
# 

# 
# Surv(surv.mq$day, surv.mq$status)[1:10]
# s1 <- survfit(Surv(day,status) ~ 1, data = surv.mq)
# str(s1)
# 
# survfit2(Surv(day,status) ~ 1, data = surv.mq) %>% 
#   ggsurvfit() +
#   labs(
#     x = "Days",
#     y = "Overall survival probability"
#   )+ 
#   add_confidence_interval()
```

## Subsetting some data for mesocosm metadata sheet

Also doing by mesocosm id instead of treatment like above

```{r}
pupae.wts.meso.sum <- pupae.wts %>%
  dplyr::group_by(Food_type,Microbe_treatment,Mesocosm_id,Exp_day) %>%
  dplyr::summarize(biomass_sum=sum(Dry_weightAVG)) %>%
  dplyr::mutate(biomass_cumsum=cumsum(biomass_sum))

#write.csv(pupae.wts.meso.sum,file="pupae.wts.mesocosm.summed.csv")
```

# Statz

## Libraries

```{r}
library(plyr)
library(dplyr)
library(tidyr)
```

## Data frame setups

```{r}
pupae.wts.exp <- subset(pupae.wts,Microbe_treatment=="ECO"|Microbe_treatment=="EMO"|Microbe_treatment=="MMO"|Microbe_treatment=="MEM"|Microbe_treatment=="ALL")
table(pupae.wts.exp$Microbe_treatment)

aggregate(Dry_weightAVG~Exp_day+Mesocosm_id, data=pupae.wts.exp, FUN=sum)

biomass <- with(merge(expand.grid(Exp_day=8:40, Mesocosm_id=unique(pupae.wts.exp$Mesocosm_id), value = 0), pupae.wts.exp, all = TRUE), aggregate(Dry_weightAVG, by=list(Exp_day, Mesocosm_id), FUN=sum))
biomass[is.na(biomass)] <- 0

names(biomass)[1]<-paste("day")
names(biomass)[2]<-paste("mesocosmID")
names(biomass)[3]<-paste("biomass")

pupate <- with(merge(expand.grid(Exp_day=8:40,Mesocosm_id=c(unique(pupae.wts.exp$Mesocosm_id)), value=0), pupae.wts.exp, all=TRUE), aggregate(Pupated, by=(list(Exp_day, Mesocosm_id)), FUN=sum))
pupate[is.na(pupate)] <- 0

names(pupate)[1]<-paste("day")
names(pupate)[2]<-paste("mesocosmID")
names(pupate)[3]<-paste("total.pupation")

masspupated <- cbind(biomass, pupate)
##write.csv(massemerged, "massemerged.csv")

##making a copy so I don't overwrite:
tallydata <- masspupated
#tallydata <- read.csv("massemerged.csv", header=TRUE)
tallydata$diet <- ifelse(grepl("B_", tallydata$mesocosmID), "Bamboo", "Larval")

tallydata$microbe <- substr(tallydata$mesocosmID,3,5)

tallydata$period <- ifelse(tallydata$day<=10, "A", ifelse(tallydata$day<=20, "B", ifelse(tallydata$day<=30, "C", "D")))
                                    
tallydata$factorial <- paste(tallydata$diet, tallydata$microbe)
```

## Fit stat models

```{r}
#fit statistical models
##GLMMs
library(glmmTMB)

tallydata$microbe<-as.factor(tallydata$microbe)
tallydata$diet<-as.factor(tallydata$diet)
tallydata$rowid<-1:nrow(tallydata)

#compare likelihood using IC between quasipoisson, gaussian, tweedie.
summary(quasinb <- glmmTMB(biomass~diet*microbe*scale(day)+diet*microbe*scale(I(day^2))+(1|mesocosmID), data=tallydata, family=nbinom1))

summary(normie <- glmmTMB(biomass~diet*microbe*scale(day)+diet*microbe*scale(I(day^2))+(1|mesocosmID), data=tallydata, family=gaussian))

summary(tweediem <- glmmTMB(biomass~diet*microbe*scale(day)+diet*microbe*scale(I(day^2))+(1|mesocosmID), data=tallydata, family=tweedie()))

AIC(quasinb)
AIC(normie)
AIC(tweediem)
#tweedie familiy assuming compound Poisson-gamma distribution best fit to these data which makes sense given its a continuous variable with many zeros
#compound Poisson-gamma does a substantially better job at handling random intercept
###quadratic temporal mixed models
```

## Tweedie models

```{r}
summary(tweediefit1 <- glmmTMB(biomass~diet*microbe*scale(day)+diet*microbe*scale(I(day^2))+(1|mesocosmID), data=tallydata, family=tweedie()))
summary(tweediefit2 <- glmmTMB(biomass~diet*microbe+diet*scale(day)+diet*scale(I(day^2))+microbe*scale(day)+microbe*scale(I(day^2))+(1|mesocosmID), data=tallydata, family=tweedie()))
anova(tweediefit1,tweediefit2)
sigma(tweediefit1)
family_params(tweediefit1)
family_params(tweediefit2)
#p=4.969e-05, full three-order quadratic model more likely than second order model 
###global test of the central hypothesis that the
###quadratic pattern of pupal biomass generation over time
###is dependent on an interaction between detritus type and microbe treatment
```

## Ggpredict

```{r}
#newdata <- cbind(tallydata$microbe, tallydata$diet, tallydata$day)
#colnames(newdata)<- c("microbe","diet","day")
library(ggeffects)

new.colors <- c("#7D0112","#C6A947","#84D1B5","#0085A6","#1F28A2")

#visualize model
predict <- ggpredict(tweediefit1, terms=c("microbe", "diet", "day[all]"), type="fixed")

predictbamboo <- subset(predict, group=="Bamboo")

predictbamboo$x <- factor(predictbamboo$x,levels=c("ECO","MMO","EMO","MEM","ALL"))
predictbamboo$facet <- as.numeric(as.character(predictbamboo$facet))

ggpred.bamboo <- ggplot(predictbamboo, aes(x=facet, y=sqrt(predicted), group=x))+
  geom_line(aes(color=x))+
  scale_x_continuous(limits=c(8,40))+
  theme_cowplot()+
  scale_color_manual(name="Microorganisms",values=new.colors,labels=c("Lab strains (LAB)","Mosquito (MOS)","Environmental (ENV)","MOS+ENV","MOS+ENV+LAB"))+
  xlab("Experimental day")+
  ggtitle("Bamboo diet")
ggpred.bamboo

predictlarval <- subset(predict, group=="Larval")

predictlarval$x <- factor(predictlarval$x,levels=c("ECO","MMO","EMO","MEM","ALL"))
predictlarval$facet <- as.numeric(as.character(predictlarval$facet))

ggpred.larval <- ggplot(predictlarval, aes(x=facet, y=sqrt(predicted), group=x))+
  geom_line(aes(color=x))+
  scale_x_continuous(limits=c(8,40))+
  theme_cowplot()+
  scale_color_manual(name="Microorganisms",values=new.colors,labels=c("Lab strains (LAB)","Mosquito (MOS)","Environmental (ENV)","MOS+ENV","MOS+ENV+LAB"))+
  xlab("Experimental day")+
  ggtitle("Standard diet")
ggpred.larval

library("ggpubr")
ggarrange(ggpred.larval,ggpred.bamboo,labels=c("(a)","(b)"),common.legend=T,legend="right")
#ggsave("facet.diet.rate.pdf",width=8,height=3)
#ggsave("facet.diet.rate.png",width=8,height=3)
```

## Stats subset by diet

### Bamboo

```{r}
##diet sub-models to visualize the effects in the global model
#bamboo detritus
bamboo <- subset(tallydata, diet=="Bamboo")
summary(bamtweediefit1 <- glmmTMB(biomass~microbe*scale(day)+microbe*scale(I(day^2))+(1|mesocosmID), data=bamboo, family=tweedie()))

summary(bamtweediefit2 <- glmmTMB(biomass~microbe+scale(day)+scale(I(day^2))+(1|mesocosmID), data=bamboo, family=tweedie()))
anova(bamtweediefit1,bamtweediefit2) ##sig***

bamboo$microbe <- relevel(bamboo$microbe, ref="MMO")
summary(bamtweediefit1 <- glmmTMB(biomass~microbe*scale(day)+microbe*scale(I(day^2))+(1|mesocosmID), data=bamboo, family=tweedie()))

bamboo$microbe<-relevel(bamboo$microbe, ref="ALL")
summary(bamtweediefit1 <- glmmTMB(biomass~microbe*scale(day)+microbe*scale(I(day^2))+(1|mesocosmID), data=bamboo, family=tweedie()))

bamboo$microbe<-relevel(bamboo$microbe, ref="MEM")
summary(bamtweediefit1 <- glmmTMB(biomass~microbe*scale(day)+microbe*scale(I(day^2))+(1|mesocosmID), data=bamboo, family=tweedie()))

bamboo$microbe<-relevel(bamboo$microbe, ref="EMO")
summary(bamtweediefit1 <- glmmTMB(biomass~microbe*scale(day)+microbe*scale(I(day^2))+(1|mesocosmID), data=bamboo, family=tweedie()))
```

### Larval food

```{r}
larval <- subset(tallydata, diet=="Larval")

larval$microbe <- relevel(larval$microbe, ref="MMO")

summary(lartweediefit1 <- glmmTMB(biomass~microbe*scale(day)+microbe*scale(I(day^2))+(1|mesocosmID), data=larval, family=tweedie()))

summary(lartweediefit2 <- glmmTMB(biomass~microbe+scale(day)+scale(I(day^2))+(1|mesocosmID), data=larval, family=tweedie()))

anova(lartweediefit1,lartweediefit2)

larval$microbe <- relevel(larval$microbe, ref="EMO")

summary(lartweediefit1 <- glmmTMB(biomass~microbe*scale(day)+microbe*scale(I(day^2))+(1|mesocosmID), data=larval, family=tweedie()))

larval$microbe <- relevel(larval$microbe, ref="MEM")

summary(lartweediefit1 <- glmmTMB(biomass~microbe*scale(day)+microbe*scale(I(day^2))+(1|mesocosmID), data=larval, family=tweedie()))
```

# Version control

```{r}
sessionInfo()
```
