---
title: "SSU real script"
date: "`r Sys.Date()`"
output:
 rmarkdown::html_document:
    theme: cerulean
    toc: yes
    toc_float: yes
    highlight: haddock
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Setup
  
## Libraries
  
```{r}
# Install the required package
#install.packages("readr")
# Load the installed Package
library(readr)
library("stringr")
library("phyloseq")
```

## Data in

```{r}
setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/18S_fungi")

counts.tab <- read.csv("bamboomesos_fungi_counts.csv", row.names=1, check.names=FALSE)

tax.tab <- read.csv("bamboomesos_fungi_taxa_wseqs.csv", row.names=1, check.names=FALSE)

metadata <- read.csv("../bamboomesos_metadata_plusmgs.csv", row.names=1, check.names=FALSE)

row.names(tax.tab)==row.names(counts.tab)

#list.bad <- c("P1-NTC2","P1-NTC3","P2-NTC1","P2-NTC2")
#metadata2 <- metadata[!row.names(metadata) %in% list.bad,]

ps.object <- phyloseq(otu_table(as.matrix(counts.tab),taxa_are_rows=T),
                      tax_table(as.matrix(tax.tab)),
                      sample_data(metadata))
ps.object #900 taxa, 176 samples
```

# Cleaning

## Filtering non-fungi

```{r}
#ps.mito <- subset_taxa(ps.object, Family=="Mitochondria")
#ps.mito #0 taxa to remove (no mitochondria)
#ps.chlor <- subset_taxa(ps.object, Order=="Chloroplast")
#ps.chlor #0 taxa to remove
ps.noteuk <- subset_taxa(ps.object, Kingdom!="Eukaryota")
ps.noteuk #232 taxa to remove

# ps.nomito <- subset_taxa(ps.object, Family!="Mitochondria")
# ps.nomito 
# ps.nochlor <- subset_taxa(ps.nomito, Order!="Chloroplast")
# ps.nochlor
ps.clean <- subset_taxa(ps.object, Kingdom=="Eukaryota")
ps.clean #668 taxa
Taxonomy <- as.data.frame(ps.clean@tax_table)
phyla.list <- unique(Taxonomy$Phylum)
phyla.list

### Keeping only Fungi
list.keep <- c("Basidiomycota","Ascomycota","Blastocladiomycota","Chytridiomycota","Cryptomycota","LKM15","Kingdom_Eukaryota")
#need to assign the Kingdom_Eukaryota taxa that we're keeping to their BLAST IDs
ps.fungi <- subset_taxa(ps.clean, Phylum %in% list.keep)
ps.fungi #408 taxa
blast.list <- c("Kingdom_Eukaryota")
ps.blast <- subset_taxa(ps.clean, Phylum %in% blast.list)
ps.blast #12 taxa
ps.blast@tax_table

ps.final <- subset_taxa(ps.fungi, !(asv.ids %in% c("ASV216", "ASV611")))
ps.final #406 taxa, removed two that did not come back as Fungi in BLAST

##Renee blasted the following, so just need to replace the taxa info 
taxa.blast12 <- read.csv("bamboomesos_fungi_taxa_wseqs_blast12.csv",row.names=1)
ps.final@tax_table <- tax_table(as.matrix(taxa.blast12))
ps.final
```

## Decontam

```{r}
# if (!requireNamespace("BiocManager", quietly = TRUE))
#   install.packages("BiocManager")
# 
# BiocManager::install("decontam")
library(phyloseq); packageVersion("phyloseq")
library(ggplot2); packageVersion("ggplot2")
#install.packages("decontam")
library(decontam); packageVersion("decontam")

head(sample_data(ps.final))
df <- as.data.frame(sample_data(ps.final))
df$LibrarySize <- sample_sums(ps.final)
df <- df[order(df$LibrarySize),]
df$Index <- seq(nrow(df))
ggplot(data=df, aes(x=Index, y=LibrarySize, color=Sample_type)) + geom_point()

sample_data(ps.final)$is.neg <- sample_data(ps.final)$Sample_type == "Negative"
otu_table(ps.final)[is.na(otu_table(ps.final))] <- 0 #replacing all the NA with 0
contamdf.prev05 <- isContaminant(ps.final, method="prevalence", neg="is.neg", threshold=0.5)
table(contamdf.prev05$contaminant) #398 False, 8 True
ps.pa <- transform_sample_counts(ps.final, function(abund) 1*(abund>0))
ps.pa.neg <- prune_samples(sample_data(ps.pa)$Sample_type == "Negative", ps.pa)
ps.pa.pos <- prune_samples(sample_data(ps.pa)$Sample_type != "Negative", ps.pa)
df.pa <- data.frame(pa.pos=taxa_sums(ps.pa.pos), pa.neg=taxa_sums(ps.pa.neg),
                    contaminant=contamdf.prev05$contaminant)
ggplot(data=df.pa, aes(x=pa.neg, y=pa.pos, color=contaminant)) + geom_point() +
  xlab("Prevalence (Negative Controls)") + ylab("Prevalence (True Samples)")

##Getting rid of contaminants
ps.final.clean <- prune_taxa(!contamdf.prev05$contaminant,ps.final)
ps.final.clean #398
ps.final.clean2 <- prune_samples(sample_sums(ps.final.clean)!=0,ps.final.clean)
ps.final.clean2
##don't use these samples:
#ps.fin <- subset_samples(ps.final.clean2,Exp_day!=20)
ps.fin <- subset_samples(ps.final.clean2, is.na(Exp_day) | Exp_day != 20)
ps.fin
ps.fin@sam_data$Sample_type

##raw reads for trimming levels below
seqtab.fin.sums <- data.frame(t(ps.fin@otu_table))
seqtab.fin.sums$sums <- rowSums(seqtab.fin.sums)
sample.sums.df <- seqtab.fin.sums[,"sums", drop=FALSE]
```

## Extra look at positive & neg controls

```{r}
ps.fin.pos1 <- subset_samples(ps.fin,Sample_type=="Positive")

ps.fin.pos <- prune_taxa(taxa_sums(ps.fin.pos1)!=0,ps.fin.pos1)
otu.pos <- data.frame(ps.fin.pos@otu_table)
#ASV5 and 24 are the true positives
#background reads is highest 20 in 62265 = 0.0003212077
#or 13 in 37764 = 0.0003442432
#will trim 0.0005, which is <0.05%
```

## Removing singletons

```{r}
# Convert to presence/absence
ps_pa <- transform_sample_counts(ps.final, function(abund) 1*(abund>0))

# Count samples per taxon
sample_counts_per_taxon <- taxa_sums(ps_pa)

# Identify taxa present in more than one sample
taxa_multi_sample <- names(sample_counts_per_taxon[sample_counts_per_taxon > 1])

# Filter the original object
ps.nosing <- prune_taxa(taxa_multi_sample, ps.fin)
ps.nosing
```

## Trimming <0.05% abundance per sample 

Decided on this threshold by examining the results of the positive control reads

```{r}
seqtab.nosing <- data.frame(t(ps.nosing@otu_table))

##checking the sample sums line up with the current otu table
row.names(seqtab.nosing)==row.names(sample.sums.df)

for (i in 1:nrow(seqtab.nosing)) {
  #row_sum <- sum(seqtab.trimmed[i,])
  threshold <- sample.sums.df[i,] * 0.0005
  #print(threshold)
  seqtab.nosing[i,seqtab.nosing[i,] <= threshold] <- 0
}

tail(sort(colSums(seqtab.nosing),decreasing=T),500)
##remove 0s
seqtab.nosing2 <- seqtab.nosing[,colSums(seqtab.nosing)>0]
##52 ASVs
tail(sort(colSums(seqtab.nosing2),decreasing=T),500)

##put new seqtab back into phyloseq object
ps.fin.trim1 <- ps.nosing
ps.fin.trim1@otu_table <- otu_table(seqtab.nosing2,taxa_are_rows=F)
ps.fin.trim1
##remove missing taxa
ps.fin.trim <- prune_taxa(taxa_sums(ps.fin.trim1)!=0,ps.fin.trim1)
ps.fin.trim
##removing missing samples
# ps.fin.trim <- prune_samples(sample_sums(ps.fin.trim2)!=0,ps.fin.trim2)
# ps.fin.trim
##52 taxa, 101 samples
##saving counts table for filtered read counts
otu.trim.fin <- data.frame(ps.fin.trim@otu_table)
#write.csv(otu.trim.fin,file="otutab.filtered.csv")
```

# Bar plots

## E. coli investigations - stock

```{r}
ps.stocks1 <- subset_samples(ps.fin.trim,Mesocosm_type=="Microbe stock")
ps.stocks <- prune_taxa(taxa_sums(ps.stocks1) > 0, ps.stocks1)
ps.stocks

ps.stocks.rel <- transform_sample_counts(ps.stocks, function(x) x / sum(x))

plot_bar(ps.stocks.rel,fill="Order")
plot_bar(ps.stocks,fill="Order")
ps.stocks@otu_table

##before I removed singletons, E. coli stock had the following singletons:
# ##to check out: asvs 057, 123, 132
# 
# asv57.1 <- subset_taxa(ps.final.clean2,asv.ids=="ASV057")
# asv57 <- prune_samples(sample_sums(asv57.1)!=0,asv57.1)
# plot_bar(asv57)
# ##well that's convenient - just the E. coli stock
# asv123.1 <- subset_taxa(ps.fin,asv.ids=="ASV123")
# asv123 <- prune_samples(sample_sums(asv123.1)!=0,asv123.1)
# plot_bar(asv123)
# ##same 
# asv132.1 <- subset_taxa(ps.fin,asv.ids=="ASV132")
# asv132 <- prune_samples(sample_sums(asv132.1)!=0,asv132.1)
# plot_bar(asv132,fill="asv.ids")
# ##same
```

## Samples

```{r}
ps.eco1 <- subset_samples(ps.fin.trim, Microbe_treatment=="ECO")
ps.eco2 <- prune_taxa(taxa_sums(ps.eco1)!=0,ps.eco1)
ps.eco <- prune_samples(sample_sums(ps.eco2)!=0,ps.eco2)
ps.eco
ps.eco@sam_data$Mesocosm_treatment

plot_bar(ps.eco,fill="Genus")

plot_bar(ps.fin.trim)+
  facet_wrap(~Microbe_treatment,scales="free")
```

## Controls

```{r}
ps.no20 <- subset_samples(ps.final.clean2, Exp_day!=20)
ps.mem2 <- subset_samples(ps.no20, Microbe_treatment %in% c("MMO", "EMO", "MEM"))
ps.mem2

#install.packages("remotes")
#remotes::install_github("KarstensLab/microshades", dependencies = TRUE)
#remotes::install_github("mikemc/speedyseq")
library("speedyseq")
library("microshades")
ps.mem <- subset_samples(ps.final.clean2, Exp_day != 20)
ps.mem
ps.mem2 <- subset_samples(ps.mem, Microbe_treatment == "MMO" | Microbe_treatment == "EMO" | Microbe_treatment == "MEM")
ps.mem2
table(ps.mem2@sam_data$Treatment_notes)
mdf_prep <- prep_mdf(ps.mem2, subgroup_level = "Genus",remove_na = TRUE)

ps.mem2 <- subset_samples(ps.mem, Microbe_treatment %in% c("MMO","EMO","MEM") & Mesocosm_type=="Experiment")
ps.mem2

ps.mem.order <- tax_glom(ps.mem2, taxrank = "Order") #

# Calculate taxa sum
sort(rowSums(data.frame(ps.mem.order@otu_table)),decreasing=TRUE)
# Combine count and taxonomyTable
#top5 = cbind(as.data.frame(tax_table(ps.mem.order)[names(top5),]), Count = top5)
#top5
color_objs_GP <- create_color_dfs(mdf_prep, group_level= "Order", subgroup_level = "Genus",
                                  selected_groups = 
                                    c("Saccharomycetales", "Malasseziales", "Chytridiales", "Cantharellales") ,cvd = TRUE)

# Extract
mdf_GP <- color_objs_GP$mdf
cdf_GP <- color_objs_GP$cdf

#plot 
#table()
plot_microshades(mdf_GP, cdf_GP) +
  facet_wrap(Treatment_notes~Sample_type,scales="free", ncol=2)

otu.tab <- t(data.frame(ps.final.clean2@otu_table))
library("vegan")

otu.tab20plus <- otu.tab[rowSums(otu.tab)>20000,]

rarecurve(otu.tab20plus,step=100,cex=0.5,xlim=c(20000,100000),main=">20k")

#otu.rarecurve <- rarecurve(otu.tab,step=100,cex=0.5)

##10-20
otu.tab1020 <- otu.tab[rowSums(otu.tab)<20000 & rowSums(otu.tab)>10000,]

rarecurve(otu.tab1020,step=100,cex=0.5,xlim=c(10000,20000),main="10-20k")

##5-10
otu.tab510 <- otu.tab[rowSums(otu.tab)<10000 & rowSums(otu.tab)>5000,]

rarecurve(otu.tab510,step=100,cex=0.5,xlim=c(5000,10000),main="5-10k")

##2-5k
otu.tab25 <- otu.tab[rowSums(otu.tab)<5000 & rowSums(otu.tab)>2000,]

rarecurve(otu.tab25,step=100,cex=0.5,xlim=c(2000,5000),main="2-5k")

##0-2k
otu.tab25 <- otu.tab[rowSums(otu.tab)<2000 & rowSums(otu.tab)>0,]

rarecurve(otu.tab25,step=100,cex=0.5,xlim=c(0,2000),main="0-2k")

otu.rareslope<- rareslope(otu.tab,sample=1000)
otu.rareslope[otu.rareslope>0.001]

otu.rarecurve <- rarecurve(otu.tab,step=10,cex=0.5,xlim=c(0,200))

saturation_check <- sapply(otu.rarecurve, function(x) {
  length_x <- length(x)
  if (length_x > 2) {
    slope <- (x[length_x] - x[length_x - 1]) / 100  # Change in species per step
    return(slope)
  } else {
    return(NA)
  }
})

saturated_samples <- names(saturation_check[saturation_check < 0.01])  # Threshold for saturation
saturated_samples


##Negative controls
negative_control <- subset_samples(ps.final.clean2, Sample_type == "Negative")
negative_control #406 taxa
taxa_sums(negative_control)
negative_control2 <- prune_taxa(taxa_sums(negative_control) !=0, negative_control)
negative_control2
plot_bar(negative_control2, fill="Genus")

##Positive Controls
positive_control <- subset_samples(ps.final.clean2, Sample_type == "Positive")
positive_control #406 taxa
positive_control2 <- prune_taxa(taxa_sums(positive_control) !=0, positive_control)
positive_control2 #5 taxa
plot_bar(positive_control2, fill="Genus")
```

## Microbe stocks

```{r}
ps.clean.rel  = transform_sample_counts(ps.final.clean2, function(x) x / sum(x) )
ps.clean.rel
microbe_stock <- subset_samples(ps.clean.rel,Sample_type == "Microbe Stock")
microbe_stock #398 taxa
microbe_stock2 <- prune_taxa(taxa_sums(microbe_stock)!=0,microbe_stock)
microbe_stock2 #64 taxa
plot_bar(microbe_stock2, fill="Phylum")

custom_palette <- c("Basidiomycota" = "#56B4E9", "Kingdom_Eukaryota" = "#E69F00", "Ascomycota" = "#009E73", 
                    "Cryptomycota" = "#cc79A7", "Blastocladiomycota" = "#F0E442", "Chytridiomycota" = "#999999","LKM15" = "#0072B2")
new_x_labels <- c("M1" = "E.coli Treatment", "M2" = "Mosquito Microbes", 
                  "M3" = "Environmental Microbes")
plot_bar(microbe_stock2, fill="Phylum") +
  scale_fill_manual(values=custom_palette) +
  scale_x_discrete(labels = new_x_labels) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),plot.title = element_text(size = 14, face = "bold", hjust = 0.5, 
                                                                                     margin = margin(t = 10, b = 20))) +
  xlab("Microbe Treatments") +
  ggtitle("Relative abundance of fungal taxa across different microbe treatments")

##different version with the black bars gone
microbe_stock_alt <- tax_glom(microbe_stock2,"Phylum")

plot_bar(microbe_stock_alt, fill="Phylum") +
  scale_fill_manual(values=custom_palette) +
  scale_x_discrete(labels = new_x_labels) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5),plot.title = element_text(size = 14, face = "bold", hjust = 0.5, 
                                                                                     margin = margin(t = 10, b = 20))) +
  xlab("Microbe Treatments") +
  ggtitle("Relative abundance of fungal taxa across different microbe treatments")
```

## Experiment samples

```{r}
ps.exp.samps <- subset_samples(ps.clean.rel,Exp_day!=20 & Exp_day!=0 & Microbe_treatment!= "AHK" & Mesocosm_type=="Experiment")

##just larvae
ps.exp.lar1 <- subset_samples(ps.exp.samps,Sample_type=="Larvae")
ps.exp.lar <- prune_taxa(taxa_sums(ps.exp.lar1)>0,ps.exp.lar1)
ps.exp.lar

plot_bar(ps.exp.lar, fill="Phylum")+
  facet_wrap(Treatment_notes~Sample_type,scales="free")+
  scale_fill_manual(values=custom_palette)
  
##just water
ps.exp.wat1 <- subset_samples(ps.exp.samps,Sample_type=="Water")
ps.exp.wat <- prune_taxa(taxa_sums(ps.exp.wat1)>0,ps.exp.wat1)
ps.exp.wat

plot_bar(ps.exp.wat, fill="Phylum")+
  facet_wrap(Treatment_notes~Sample_type,scales="free")+
  scale_fill_manual(values=custom_palette)
  
```

# Alpha Diversity

## Water & larvae samples

```{r}
# theme_set(theme_bw())
# pal = "Set1"
# scale_colour_discrete <- function(palname=pal, ...){scale_colour_brewer(palette=palname, ...)}
# scale_fill_discrete <- function(palname=pal, ...){scale_fill_brewer(palette=palname, ...)}

plot_richness(ps.final.clean2,x="Sample_type", measures=c("Chao1", "Shannon"))

df.all <- data.frame(estimate_richness(ps.final.clean2, split=TRUE, measures=c("Shannon","InvSimpson", "Observed","Chao1")))
df.all$Short_label <- rownames(df.all)
metadata2$Short_label <- rownames(metadata2)
df.all.div <- merge(df.all,metadata2,by="Short_label")
df.div.stocks <- subset(df.all.div,Mesocosm_type!="Additional control")
ggplot(df.div.stocks,aes(x=Microbe_treatment,y=Observed))+
  facet_wrap(Sample_type+Food_type~Mesocosm_type,scales="free")+
  #scale_color_manual(values=stock.colors)+
  geom_boxplot()

## Box Plots Microbe Treatments vs Larvae/water vs Diet type
#install.packages("RColorBrewer")  # For better colors
#library(ggplot2)
#library(RColorBrewer)
theme_set(theme_bw())
#pal ="Set1"
cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
#scale_colour_discrete <- function(palname=pal, ...){scale_colour_brewer(palette=palname, ...)}
#scale_fill_discrete <- function(palname=pal, ...){scale_fill_brewer(palette=palname, ...)}
df.div.new <- subset(df.div.stocks, 
                        !Microbe_treatment %in% c("Microbe stock", "Negative","AHK") & 
                          !Food_type %in% c("Hatching flask", "Lab water", "Microbe stock", "Negative"))

ggplot(df.div.new, aes(x = Microbe_treatment, y = Observed, fill = Microbe_treatment)) +
  geom_boxplot() +  
  facet_grid(Sample_type ~ Food_type, labeller = labeller(Food_type = c("Larval food" = "Lab Diet", "Bamboo" = "Bamboo Powder"))) +  # Splits by Water/Larvae and Diet
  scale_fill_manual(values=cbPalette, labels = c("All treatments","E. coli only","Environmental microbes only","Mosquito & Environmental microbes","Mosquito microbes only")) + #Color Blind Friendly palette
  xlab("Microbe Treatments") +
  theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5,
                                  margin = margin(t = 20, b = 20)))+
  ggtitle("Observed fungal taxa across microbe treatments and diet types for larvae and water samples")
  #scale_fill_brewer(palette = "Set2")  # Pastel colors


```

## Microbe stocks plot

```{r}
df.div.ms <- subset(df.div.stocks, 
                     Mesocosm_treatment %in% c("Microbe_stock_ecoli","Microbe_stock_mosq","Microbe_stock_brom"))
ggplot(df.div.ms, aes(x = Mesocosm_treatment, y = Observed, fill = Mesocosm_treatment)) +
  geom_bar(stat="identity") +  
  #facet_grid(Sample_type ~ Food_type) +  # Splits by Water/Larvae and Diet
  theme(legend.position="none")+
  scale_fill_manual(values=c("#009E73","#56B4E9","#0072B2"))

```

## Stats (mosquitoes)

```{r}
df.div.mos <- subset(df.div.new, 
                     Sample_type == "Larvae" & Microbe_treatment != "AHK")
##Histogram
hist(df.div.mos$Observed) #log was worse for Observed
##ANOVA even though not normally distributed (first assumption violated)
two.way <- aov(Observed ~ Microbe_treatment + Food_type, data=df.div.mos)
summary(two.way) #food type is significant, microbe treatment not
interaction <- aov(Observed ~ Microbe_treatment * Food_type, data=df.div.mos)
summary(interaction)
#install.packages('car')
library('car')
library('ggplot2')
##Levene Test for Homogeneity in Variance
leveneTest(Observed ~ Microbe_treatment * Food_type, data=df.div.mos) #variance is not too bad (bad variance would produce a significant P value)
ggplot(df.div.mos, aes(x = Microbe_treatment, y = Observed, fill = Microbe_treatment)) +
  geom_boxplot() 
ggplot(df.div.mos, aes(x = Food_type, y = Observed, fill = Food_type)) +
  geom_boxplot() 
```

## More advanced stats

### Richness - larvae

```{r}
library('stats')

glm <- glm(Observed ~ Microbe_treatment * Food_type, data=df.div.mos, family = poisson)
summary(glm)
glm.normal <- glm(Observed ~ Microbe_treatment * Food_type, data=df.div.mos) #without family=poisson, assumes gaussian distribution
summary(glm.normal)
AIC(glm,glm.normal) #compares the two and tells which fits the data better, lower AIC fits better
#glm has AIC 132.9 glm.normal has AIC 138.8
#install.packages('MASS')
library('MASS')
glm.nb <- glm.nb(Observed ~ Microbe_treatment * Food_type, data=df.div.mos) #checking negative binomial
summary(glm.nb)
AIC(glm,glm.normal,glm.nb) #poisson is still best fit, nb is better than normal

glm <- glm(Observed ~ Microbe_treatment * Food_type, data=df.div.mos, family = poisson)
summary(glm)
glm.plus <- glm(Observed ~ Microbe_treatment + Food_type, data=df.div.mos, family = poisson)
summary(glm.plus)
AIC(glm,glm.plus)
anova(glm,glm.plus) #glm.plus is final model because no significant interaction (*)

glm.plus.nomicrobes <- glm(Observed ~ Food_type, data=df.div.mos, family = poisson)
AIC(glm.plus,glm.plus.nomicrobes)
anova(glm.plus.nomicrobes,glm.plus) #P value = 0.1256 Not significant

glm.plus.nofood <- glm(Observed ~ Microbe_treatment, data=df.div.mos, family = poisson)
AIC(glm.plus,glm.plus.nofood)
anova(glm.plus,glm.plus.nofood) #P value = 0.001969 Significant
```

### Shannon - larvae

```{r}
df.div.mos.mem <- subset(df.div.mos,Microbe_treatment %in% c("EMO","MMO","MEM"))
##Repeat for Shannon for larvae
##include 239-249 lines for shannon

glm.shan <- glm(Shannon ~ Microbe_treatment * Food_type, data=df.div.mos.mem, family = poisson)
summary(glm.shan)
glm.shan.normal <- glm(Shannon ~ Microbe_treatment * Food_type, data=df.div.mos.mem) #without family=poisson, assumes gaussian distribution
summary(glm.shan.normal)
AIC(glm.shan,glm.shan.normal) #compares the two and tells which fits the data better, lower AIC fits better
#glm.shan has AIC Inf glm.shan.normal has AIC 30.43396
#install.packages('MASS')
library('MASS')
glm.shan.nb <- glm.nb(Shannon ~ Microbe_treatment * Food_type, data=df.div.mos.mem) #checking negative binomial
summary(glm.shan.nb)
AIC(glm.shan,glm.shan.normal,glm.shan.nb) #lowest AIC is glm.shan.normal at 30.43396

glm.shan.plus <- glm(Shannon ~ Microbe_treatment + Food_type, data = df.div.mos.mem)
summary(glm.shan.plus)
AIC(glm.shan.normal,glm.shan.plus) #glm.shan.plus is final model for Shannon, no significant interaction again
anova(glm.shan.normal,glm.shan.plus)#P value = 0.275

glm.shan.plus.nomic <- glm(Shannon ~ Food_type, data=df.div.mos.mem)
AIC(glm.shan.plus,glm.shan.plus.nomic)
anova(glm.shan.plus,glm.shan.plus.nomic) #P value = 0.153

glm.plus.shan.nofood  <- glm(Shannon ~ Microbe_treatment, data=df.div.mos.mem)
AIC(glm.shan.plus, glm.plus.shan.nofood) #glm.plus.shan.nofood
anova(glm.shan.plus,glm.plus.shan.nofood) #P value 0.2161

glm.shan.plus.em <- emmeans(glm.shan.plus, ~Food_type+Microbe_treatment)

summary(glm.shan.plus.em)

glm.shan.plus.em.cld <- cld(glm.shan.plus.em,Letters=letters)
summary(glm.shan.plus.em.cld)
```

### Richness - water

```{r}
df.div.wat <- subset(df.div.new, 
                     Sample_type == "Water" & Microbe_treatment != "AHK" & Exp_day!=20)

##code here
```

### Shannon - water

```{r}
df.div.wat.mem.mem <- subset(df.div.wat.mem,Microbe_treatment %in% c("EMO","MMO","MEM"))
##Repeat for Shannon for larvae
##include 239-249 lines for shannon

glm.shan <- glm(Shannon ~ Microbe_treatment * Food_type, data=df.div.wat.mem.mem, family = poisson)
summary(glm.shan)
glm.shan.normal <- glm(Shannon ~ Microbe_treatment * Food_type, data=df.div.wat.mem) #without family=poisson, assumes gaussian distribution
summary(glm.shan.normal)
AIC(glm.shan,glm.shan.normal) #compares the two and tells which fits the data better, lower AIC fits better
#glm.shan has AIC Inf glm.shan.normal has AIC 30.43396
#install.packages('MASS')
library('MASS')
glm.shan.nb <- glm.nb(Shannon ~ Microbe_treatment * Food_type, data=df.div.wat.mem) #checking negative binomial
summary(glm.shan.nb)
AIC(glm.shan,glm.shan.normal,glm.shan.nb) #lowest AIC is glm.shan.normal at 30.43396

glm.shan.plus <- glm(Shannon ~ Microbe_treatment + Food_type, data = df.div.wat.mem)
summary(glm.shan.plus)
AIC(glm.shan.normal,glm.shan.plus) #glm.shan.plus is final model for Shannon, no significant interaction again
anova(glm.shan.normal,glm.shan.plus)#P value = 0.275

glm.shan.plus.nomic <- glm(Shannon ~ Food_type, data=df.div.wat.mem)
AIC(glm.shan.plus,glm.shan.plus.nomic)
anova(glm.shan.plus,glm.shan.plus.nomic) #P value = 0.153

glm.plus.shan.nofood  <- glm(Shannon ~ Microbe_treatment, data=df.div.wat.mem)
AIC(glm.shan.plus, glm.plus.shan.nofood) #glm.plus.shan.nofood
anova(glm.shan.plus,glm.plus.shan.nofood) #P value 0.2161

glm.shan.plus.em <- emmeans(glm.shan.plus, ~Microbe_treatment)

summary(glm.shan.plus.em)

pairs(glm.shan.plus.em)

glm.shan.plus.em.cld <- cld(glm.shan.plus.em,Letters=letters)
summary(glm.shan.plus.em.cld)



library("")
simulateResiduals(glm.shan.plus)
```

# Fitness from the treatments

## Boxplot

```{r}
ggplot(data=df.div.wat,aes(x=Microbe_treatment,y=Biomass_day40,fill=Microbe_treatment))+
  geom_boxplot()+
  scale_fill_manual(values=cbPalette)+
  facet_wrap(~Food_type)
```

### Stats

```{r}
library("emmeans")
library("multcomp")

metadat.fit <- subset(metadata2,Mesocosm_type=="Experiment"&Sample_type=="Water"&Microbe_treatment!="AHK"&Exp_day!=20)

##just need 1 biomass number per mesocosm
metadat.fit$Mesocosm_id

hist(metadat.fit$Biomass_day40)

ggplot(metadat.fit,aes(x=Microbe_treatment,y=Biomass_day40))+
  geom_point()+
  facet_wrap(~Food_type)

glm.biomass <- glm(Biomass_day40 ~ Microbe_treatment * Food_type, data=metadat.fit)
summary(glm.biomass)

glm.biomass.plus <- glm(Biomass_day40 ~ Microbe_treatment + Food_type, data=metadat.fit)
summary(glm.biomass.plus)

anova(glm.biomass,glm.biomass.plus) #sig interaction between microbe treatment and food type - affects output biomass

```

## Alpha diversity correlation with fitness?

### Larvae

```{r}
library("glmmTMB")

corr.lar.int <- glmmTMB(Biomass_day40~Observed*Food_type+(1|Microbe_treatment),data=df.div.mos)
summary(corr.lar.int)

corr.lar.noint <- glmmTMB(Biomass_day40~Observed+Food_type+(1|Microbe_treatment),data=df.div.mos)
summary(corr.lar.noint) #just food type which we already knew about

anova(corr.lar.int,corr.lar.noint) #not sig

ggplot(df.div.mos,aes(x=Observed,y=Biomass_day40, color=Food_type,shape=Microbe_treatment))+
  geom_point()+
  geom_smooth(method="lm",aes(group=Food_type))
```

### Water

```{r}
corr.wat.int <- glmmTMB(Biomass_day40~Observed*Food_type+(1|Microbe_treatment),data=df.div.wat)
summary(corr.wat.int)

corr.wat.noint <- glmmTMB(Biomass_day40~Observed+Food_type+(1|Microbe_treatment),data=df.div.wat)
summary(corr.wat.noint) #just food type which we already knew about

anova(corr.wat.int,corr.wat.noint) #not sig

ggplot(df.div.wat,aes(x=Observed,y=Biomass_day40, color=Food_type,shape=Microbe_treatment))+
  geom_point()+
  scale_color_manual(values=c())
  geom_smooth(method="lm",aes(group=Food_type))
```

# PCoA/NMDS

```{r}
table(ps.clean.rel@sam_data$Sample_type)
table(ps.clean.rel@sam_data$Exp_day)

ps.exp.samps <- subset_samples(ps.clean.rel,Exp_day!=20 & Exp_day!=0 & Microbe_treatment!= "AHK" & Mesocosm_type=="Experiment")
ps.exp.samps

table(ps.exp.samps@sam_data$Sample_type)
table(ps.exp.samps@sam_data$Microbe_treatment)

ord.exp <- ordinate(ps.exp.samps, "NMDS", "bray")

plot_ordination(ps.exp.samps,ord.exp,color="Sample_type",label="Microbe_treatment")

ord.exp <- ordinate(ps.exp.samps, "NMDS", "jaccard")

plot_ordination(ps.exp.samps,ord.exp,color="Sample_type",label="Microbe_treatment")

ord.exp <- ordinate(ps.exp.samps, "PCoA", "bray")

library("ggpubr")

p1 <- plot_ordination(ps.exp.samps,ord.exp,color="Food_type",shape="Microbe_treatment")

p2 <- plot_ordination(ps.exp.samps,ord.exp,color="Food_type",shape="Sample_type")

ggarrange(p1,p2)
```

## Community distance stats 

```{r}
library('vegan')

dist.bray <- phyloseq::distance(ps.exp.samps, method = "bray")

sampledf <- data.frame(sample_data(ps.exp.samps))

adonis2(dist.bray ~ Sample_type+Microbe_treatment+Food_type, data = sampledf,by="margin")

adonis2(dist.bray ~ Sample_type+Microbe_treatment*Food_type, data = sampledf,by="margin")

##significant interaction between microbe treatment and food type, no effect of sample type (meaning no differences between larvae & water communities which is very different from what we see in bacteria which is kind of cool)

```
