---
title: "bamboomesos_controlchecks"
author: "NGK"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

# Setup

```{r}
library(MicEco)
library("ggplot2")
library("phyloseq")

setwd("~/nicolagk@hawaii.edu - Google Drive/My Drive/Bamboo_mesos/Bamboo_mesos/16S_bacteria/01.process_asvs/control_checks")
```

## Data

```{r}
##ps object with all the steps except the trimming one
ps.clean.notrim <- readRDS("../bamboomesos_ps.lulu.clean.notrim.decontam.rds")
##ps object with the trimming
ps.clean.trim <- readRDS("../bamboomesos_ps.lulu.clean.decontam.rds")

##seeing if raw reads line up with expectations
samdf.clean.trim <- ps.clean.trim@sam_data
samdf.clean.trim.no20 <- subset(samdf.clean.trim,Exp_day!=20)

ggplot(samdf.clean.trim.no20,aes(x=Short_label,y=lib_size_clean,color=Microbe_treatment))+
  geom_point()+
  facet_wrap(~Microbe_treatment,scales="free_x")
```

# Microbe stocks

I looked at the ones that were trimmed <0.1% abundance before, but there are a ton that are unclear so checking less stringently which ones might be in MOS, ENV, or both

```{r}
ps.stocks1 <- subset_samples(ps.clean.trim,Sample_type=="Microbe Stock")
ps.stocks <- prune_taxa(taxa_sums(ps.stocks1)!=0,ps.stocks1)
ps.stocks

ps_venn(ps.stocks,group="Microbe_treatment",quantities=TRUE,main="ASVs per microbial stock",labels=c("LAB","ENV","MOS"),fills = list(fill = c("#7D0112", "#84D1B5","#C6A947"), alpha = 0.5))
ps_venn(ps.stocks,group="Microbe_treatment",plot=F)

#grid.draw(venn.stocks)

ps.stocks.notrim1 <- subset_samples(ps.clean.notrim,Sample_type=="Microbe Stock")
ps.stocks.notrim <- prune_taxa(taxa_sums(ps.stocks.notrim1)!=0,ps.stocks.notrim1)
ps.stocks.notrim

ps_venn(ps.stocks.notrim,group="Microbe_treatment",quantities=TRUE,main="ASVs per microbial stock",labels=c("LAB","ENV","MOS"),fills = list(fill = c("#7D0112", "#84D1B5","#C6A947"), alpha = 0.5))

#grid.draw(venn.stocks.notrim)
stocks.notrim <- ps_venn(ps.stocks.notrim,group="Microbe_treatment",quantities=TRUE,main="ASVs per microbial stock",labels=c("LAB","ENV","MOS"),fills = list(fill = c("#7D0112", "#84D1B5","#C6A947"), alpha = 0.5),plot=F)

stocks.notrim[["E. coli"]]
##2 is like in everything in the real samples & very little in the stocks, plus looks like it came from lab water :(
envonly.untrim <- stocks.notrim[["Env. microbes"]]
mosonly.untrim <- stocks.notrim[["Mosquito microbes"]]

##saving to add info to taxa file, which I will do manually in Excel
#write.table(mosonly.untrim,"mosonly.txt", sep="\t",col.names = FALSE, row.names = FALSE,quote=F)
#write.table(envonly.untrim,"envonly.txt", sep="\t",col.names = FALSE, row.names = FALSE,quote=F)
#envmos.untrim <- stocks.notrim

##others to check out:
#stocks.notrim[["E. coli__Env. microbes__Mosquito microbes"]]
# [1] "ASV0003"  "ASV0007"
#stocks.notrim[["E. coli__Env. microbes"]]
# [1] "ASV0005" "ASV0008" "ASV0009" "ASV0132"
#stocks.notrim[["E. coli__Mosquito microbes"]]
# [1] "ASV0001" = definitely from E. coli
#stocks.notrim[["Env. microbes__Mosquito microbes"]]
# [1] "ASV0006" "ASV0053" "ASV0136" "ASV0149" "ASV0525" "ASV0989" "ASV1146"

##asv0003
ps.trim.asv3.1 <- subset_taxa(ps.clean.trim, id=="ASV0003")
ps.trim.asv3 <- prune_taxa(taxa_sums(ps.trim.asv3.1)!=0,ps.trim.asv3.1)

plot_bar(ps.trim.asv3,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")
##hard to tell - will mark as both

##asv0007
ps.trim.asv7.1 <- subset_taxa(ps.clean.trim, id=="ASV0007")
ps.trim.asv7 <- prune_taxa(taxa_sums(ps.trim.asv7.1)!=0,ps.trim.asv7.1)
##hard to tell - will mark as both

plot_bar(ps.trim.asv7,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##asv0005 = clearly ENV
ps.trim.asv5.1 <- subset_taxa(ps.clean.trim, id=="ASV0008")
ps.trim.asv5 <- prune_taxa(taxa_sums(ps.trim.asv5.1)!=0,ps.trim.asv5.1)

plot_bar(ps.trim.asv5,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")
```

# Checking where microbes came from

```{r}
otu.trim <- data.frame(ps.clean.trim@otu_table)
otu.notrim <- data.frame(ps.clean.notrim@otu_table)

##sample that aren't matching up between them:
#row.names(otu.trim) %in% row.names(otu.notrim)
what <- otu.trim[!row.names(otu.trim) %in% row.names(otu.notrim),]
what2 <- otu.notrim[!row.names(otu.notrim) %in% row.names(otu.trim),]

rowSums(what)
rowSums(what2)
##just negatives, and WB39 which is a day 20 sample
```

## Hatching flask & lab water

```{r}
##which ones are in the hatching flasks - trimmed data
# - Hatching flask water
# - Hatching flask L1s
# - Lab water 

ps.base1 <- subset_samples(ps.clean.trim,Mesocosm_treatment=="Hatch_flask" | Mesocosm_treatment=="Lab_water")

ps.base <- prune_taxa(taxa_sums(ps.base1)!=0,ps.base1)
ps.base

plot_bar(ps.base,fill="id",x="Longer_name")+
  facet_wrap(~Genus,scales="free")

ps.base@otu_table
##4 taxa of note to take with a grain of salt:
##"ASV0002" "ASV0008" "ASV0018" "ASV0030" 
#Wolbachia: "ASV0227" "ASV0274"

#ASV2 = extremely convincing information that it was in the lab water - Sphingomonas
#the hatching flask water looked very clean though... confusing
#ASV0030 = in the hatching flask larvae??? 
#ASV0008 = in the hatching flask larvae - but just 80 counts total... - Novosphingobium
#ASV18 = 800+ counts in the lab water

##ASV30 = E.coli
ps.trim.asv30.1 <- subset_taxa(ps.clean.trim, id=="ASV0030")
ps.trim.asv30 <- prune_taxa(taxa_sums(ps.trim.asv30.1)!=0,ps.trim.asv30.1)

plot_bar(ps.trim.asv30,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##ASV2 = sphingomonas - in the lab water?
ps.trim.asv2.1 <- subset_taxa(ps.clean.trim, id=="ASV0002")
ps.trim.asv2 <- prune_taxa(taxa_sums(ps.trim.asv2.1)!=0,ps.trim.asv2.1)

plot_bar(ps.trim.asv2,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##asv8 = sphingomonas - in the lab water?
ps.trim.asv8.1 <- subset_taxa(ps.clean.trim, id=="ASV0008")
ps.trim.asv8 <- prune_taxa(taxa_sums(ps.trim.asv8.1)!=0,ps.trim.asv8.1)

plot_bar(ps.trim.asv8,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##asv18 = sphingomonas 
ps.trim.asv18.1 <- subset_taxa(ps.clean.trim, id=="ASV0018")
ps.trim.asv18 <- prune_taxa(taxa_sums(ps.trim.asv18.1)!=0,ps.trim.asv18.1)

plot_bar(ps.trim.asv18,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##all taxa that are in the base controls
ps.clean.basetax <- subset_taxa(ps.clean.trim, taxa_names(ps.clean.trim) %in% taxa_names(ps.base))

plot_bar(ps.clean.basetax,fill="id")+
  facet_wrap(Genus~Sample_type,scales="free")

ps.clean.wolb <- subset_taxa(ps.clean.trim,Genus=="Wolbachia")
plot_bar(ps.clean.wolb)
```

## Additional mesocosm controls

```{r}
ps.mesocontrols1 <- subset_samples(ps.clean.trim,Microbe_treatment %in% c("None_food","None_larvae"))
ps.mesocontrols1
ps.mesocontrols <- prune_taxa(taxa_sums(ps.mesocontrols1)!=0,ps.mesocontrols1)
ps.mesocontrols

plot_bar(ps.mesocontrols,fill="Genus", x="Longer_name")+
  facet_wrap(~id,scales="free")

##asv49 = Amnibacterium 
ps.trim.asv49.1 <- subset_taxa(ps.clean.trim, id=="ASV0049")
ps.trim.asv49 <- prune_taxa(taxa_sums(ps.trim.asv49.1)!=0,ps.trim.asv49.1)

plot_bar(ps.trim.asv49,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##asv68 = Amnibacterium 
ps.trim.asv68.1 <- subset_taxa(ps.clean.trim, id=="ASV0068")
ps.trim.asv68 <- prune_taxa(taxa_sums(ps.trim.asv68.1)!=0,ps.trim.asv68.1)

plot_bar(ps.trim.asv68,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##asv97 = methylo 
ps.trim.asv97.1 <- subset_taxa(ps.clean.trim, id=="ASV0097")
ps.trim.asv97 <- prune_taxa(taxa_sums(ps.trim.asv97.1)!=0,ps.trim.asv97.1)

plot_bar(ps.trim.asv97,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##asv131 - spirosoma
ps.trim.asv131.1 <- subset_taxa(ps.clean.notrim, id=="ASV0131")
ps.trim.asv131 <- prune_taxa(taxa_sums(ps.trim.asv131.1)!=0,ps.trim.asv131.1)

plot_bar(ps.trim.asv131,fill="id")+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

###same stuff as above but relative abundance
ps.mesocontrols.rel <- transform_sample_counts(ps.mesocontrols, function(x) x / sum(x))

plot_bar(ps.mesocontrols.rel,fill="id",x="Longer_name")+
  facet_wrap(~Genus,scales="free")

sample_sums(ps.mesocontrols)

##how many shared with real samples
ps.clean.trim
ps.noctrl <- subset_samples(ps.clean.trim,!Mesocosm_type %in% c("Negative","Positive"))
ps.noctrl
table(ps.noctrl@sam_data$Mesocosm_type)

ps_venn(ps.noctrl,group="Mesocosm_type",quantities=TRUE)
ps_venn(ps.noctrl,group="Mesocosm_type",quantities=TRUE,plot=F)
```

The following doesn't work anymore now that I assigned everything... 

```
##see if we accounted for all the unclear ones overlapping with known sources:
ps.noctrl.unc <- subset_taxa(ps.noctrl,likely_source_untrimmed=="unclear")
ps.noctrl.unc

table(ps.noctrl.unc@sam_data$Mesocosm_type)
ps_venn(ps.noctrl.unc,group="Mesocosm_type",quantities=TRUE)

table(ps.noctrl.unc@sam_data$Microbe_treatment)

ps.noctrl.unc.big3 <- subset_samples(ps.noctrl.unc,Microbe_treatment %in% c("ECO","EMO","MMO") & Exp_day!=20)
#table(ps.noctrl.unc.big3@sam_data$Exp_day)

ps_venn(ps.noctrl.unc.big3,group="Microbe_treatment",quantities=TRUE)

ps_venn(ps.noctrl.unc.big3,group="Microbe_treatment",quantities=TRUE,plot=F)

##checking out just eco group
ps.unc.eco1 <- subset_samples(ps.noctrl.unc,Microbe_treatment=="ECO")
ps.unc.eco <- prune_taxa(taxa_sums(ps.unc.eco1)!=0,ps.unc.eco1)
ps_venn(ps.unc.eco,group="Mesocosm_treatment",quantities=T)

plot_bar(ps.unc.eco,fill="id",x="Longer_name")+
  facet_wrap(~Genus,scales="free")
```

```{r}
##Massilia looks sus
ps.asv420 <- subset_taxa(ps.clean.trim,id=="ASV0420")
plot_bar(ps.asv420)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##Methylo 28 looks sus
ps.asv28 <- subset_taxa(ps.clean.trim,id=="ASV0028")
plot_bar(ps.asv28)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##Methylo 75 looks sus
ps.asv75 <- subset_taxa(ps.clean.trim,id=="ASV0075")
plot_bar(ps.asv75)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")
```

```
##checking out just emo & mmo group
ps.unc.envmos1 <- subset_samples(ps.noctrl.unc,Microbe_treatment %in% c("EMO","MMO") & Exp_day!=20 & Mesocosm_type=="Experiment")
ps.unc.envmos1
ps.unc.envmos <- prune_taxa(taxa_sums(ps.unc.envmos1)!=0,ps.unc.envmos1)
ps_venn(ps.unc.envmos,group="Microbe_treatment",quantities=T)
table(ps.unc.envmos@sam_data$Mesocosm_treatment)

ps_venn(ps.unc.envmos,group="Microbe_treatment",quantities=T)

venn.unc.envmos <- ps_venn(ps.unc.envmos,group="Microbe_treatment",quantities=T,plot=F)

ps.unc.envmos.8<- subset_taxa(ps.unc.envmos, id %in% c("ASV0037","ASV0057","ASV0060","ASV0065","ASV0125","ASV0130","ASV0176","ASV0181"))

plot_bar(ps.unc.envmos.8,fill="id",x="Longer_name")+
  facet_wrap(~Microbe_treatment,scales="free")
```

```{r}
ps.asv37 <- subset_taxa(ps.clean.trim,id=="ASV0037")
plot_bar(ps.asv37)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

ps.asv57 <- subset_taxa(ps.clean.trim,id=="ASV0057")
plot_bar(ps.asv57)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

ps.asv60 <- subset_taxa(ps.clean.trim,id=="ASV0060")
plot_bar(ps.asv60)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

ps.asv65 <- subset_taxa(ps.clean.trim,id=="ASV0065")
plot_bar(ps.asv65)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

ps.asv125 <- subset_taxa(ps.clean.trim,id=="ASV0125")
plot_bar(ps.asv125)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

ps.asv176 <- subset_taxa(ps.clean.trim,id=="ASV0176")
plot_bar(ps.asv176)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

ps.asv181 <- subset_taxa(ps.clean.trim,id=="ASV0181")
plot_bar(ps.asv181)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")

##based on negative controls:
ps.asv7 <- subset_taxa(ps.clean.trim,id=="ASV0007")
plot_bar(ps.asv7)+
  facet_wrap(Sample_type~Microbe_treatment,scales="free")
```

## Grabbing a ton more to add to likely ENV and MOS sources

Will do manually in excel after... Also... some of the output above won't look the same after I change the spreadsheets... oh well 

```
envonly.unc <- venn.unc.envmos[["EMO"]]
mosonly.unc <- venn.unc.envmos[["MMO"]]

##saving to add info to taxa file, which I will do manually in Excel
#write.table(mosonly.unc,"mosonly_more.txt", sep="\t",col.names = FALSE, row.names = FALSE,quote=F)
#write.table(envonly.unc,"envonly_more.txt", sep="\t",col.names = FALSE, row.names = FALSE,quote=F)
```


